# Results

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message = FALSE)
```

## Social

To discuss the social rights for women in the world these years.

To explore the details about the social rights, the heatmap is a good choice. The y axis is ordered by the average scores of countries from high to low. The x axis is ordered by the average scores of the topics.

```{r ,fig.width=12}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(parcoords)

data = read.csv("data/Gender_StatsData.csv",header = TRUE,encoding="UTF-8")
colnames(data)=c("Country","Code","Indicator_details","Indicator",as.character(1960:2020))

Code = c("CAN","USA","MEX","NIC","SLV","COL","BRA","CHL","ARG","URY","RUS","ISL","UKR","BLR","LTU","CHN","JPN","KOR","THA","MYS","COG","ZWE","TZA","NER","NGA",'AUS','NZL','PNG','WSM','FJI')
Continent = rep(c("North_America","South_America","Europe","Asia","Africa","Oceania"),rep(5,6))
data_continent = data.frame(cbind(Code,Continent))

Indicator = c("SG.APL.PSPT.EQ","SG.HLD.HEAD.EQ","SG.LOC.LIVE.EQ","SG.GET.JOBS.EQ","SG.OBT.DVRC.EQ","SG.OPN.BANK.EQ","SG.BUS.REGT.EQ","SG.CNT.SIGN.EQ","SG.HME.TRVL.EQ","SG.CTR.TRVL.EQ","SG.NGT.WORK.EQ","SG.DNG.WORK.DN.EQ","SG.IND.WORK.EQ","SG.REM.RIGT.EQ","SG.AGE.FUPN.EQ","SG.AGE.PAPN.EQ","SH.MMR.LEVE.GT","SG.LAW.ASST.AR","SG.LAW.CRDD.GR","SG.AGE.MRET.EQ","SG.ABS.PENB","SG.LEG.SXHR.EM","SG.LEG.DVAW","SG.LAW.OBHB.MR.NO")
Category = c("law","marriage","freedom","job","marriage","finance","finance","law","freedom","freedom","job","job","job","marriage","welfare","welfare","welfare","law","law","law","welfare","law","law","law")
data_cat = data.frame(cbind(Indicator,Category))

data_1 = data[data$Code%in%Code & data$Indicator%in%Indicator,c(1,2,4,25:65)]
data_2 = left_join(data_1,data_continent,by=c("Code"))
data_3 = left_join(data_2,data_cat,by=c("Indicator"))

data_4 = data_3 %>%pivot_longer(cols = starts_with("1") | starts_with("2"),names_to = "Year",values_to = "Yes_or_No")
data_c = data_4 %>% group_by(Country,Code,Continent,Category,Year) %>% summarise(Score = mean(Yes_or_No))

#data_e for reorder
data_tot_s_y = data_c[data_c$Year == 2020,] %>% group_by(Country) %>% summarise(Tot_S_Y = sum(Score))
data_tot_s_x = data_c[data_c$Year == 2020,] %>% group_by(Category) %>% summarise(Tot_S_X = sum(Score))
data_e = left_join(data_c[data_c$Year == 2020,],data_tot_s_y,by=c("Country"))
data_e = left_join(data_e,data_tot_s_x,by = c("Category"))

ggplot(data_e,aes(y=fct_reorder(Country,Tot_S_Y),x=fct_rev(fct_reorder(Category,Tot_S_X))))+
  geom_tile(aes(fill=Score)) +
  scale_fill_distiller(palette='Reds',direction = 1) +
  scale_x_discrete("Score")+
  scale_y_discrete("Country")+
  theme_classic(base_size = 16) +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank())
```

We can see that the Iceland, the Canada and the New Zealand actually both achieve 1 in 6 topics, meaning that no gender discrimination in the social rights aspect in these three countries. Compare to that, The Niger has no score 1 in every topic, meaning there is still some gender discrimination in these aspects. Especially for the marriage, the score is close to 0, meaning men have more power to dominate the marriage in Niger. Another points should be paid attention to is that the job have a very low score in Ukraine while other scores are quite high, meaning that only in the workplace, the women will be treated differently as men in Ukraine.

Moreover, there are some countries have really similar conditions in different aspects for the gender discrimination. For example, Canada and Iceland and New Zealand, United States and Mexico, Nicaragua and Lithuania, Zimbabwe and El Salvador, Fiji and Colombia, China and Belarus, Russian Federation and Papua New Guinea. We will discuss what kinds of factor contributes to the distribution of the scores like this.

Generally, women tend to have the same rights as the men in the finance aspect since only 1 country did not achieve 1 in finance. For the freedom part, only 4 countries did not achieve 1, meaning that the majorities of the countries do not keep women from doing something which men can do. However, in terms of the social welfare, the women tend to have a worse welfare guaranteed compared to men in these years with only 5 countries achieving score 1. As for law, marriage and job, different countries have different conditions and the results did not have a strong inclination to 1 or 0.

```{r fig.width=12}

ggplot(data_e, aes(Score)) +
  geom_histogram(color = "grey", 
                 fill = "lightBlue",
                 binwidth = 0.1, 
                 center = 0.05, 
                 closed = "right")+
  facet_wrap(~ Category)+
  scale_x_continuous(breaks=seq(0,1,0.1))+
  theme_bw(base_size = 16) 

#situation -> heatmap #6 aspects vs 25 countries, scores
```

Since the heatmap do not show the detailed score distribution of topics clearly. We can explore the general distribution more carefully by looking at the bar charts. We can see that for the finance part ad freedom part, the majorities of countries has 1. For the job aspect, We can see that except the score 1 and the outlier laying between 0.2 to 0.3, there are two clusters of countries distributed in the range of 0.4-0.5 and 0.7-0.8, meaning there is a gap (0.5-0.7) between the countries for the women social status in the job aspect. For law, we can see the range of the clusters are nested in the right corner except the outlier laying between 0.3 to 0.4, meaning the general scores are high in the law aspect and the difference of the condition are small between the countries. For marriage, only 3 countries are outliers, and the majority countries are landed in score 1 with approximately 7 countries having 0.6-0.7 scores. For welfare, we can see there are mainly 3 clusters, for 0.4-0.5, 0.7-0.8, 0.9-1.0 respectively. And nearly halves are in 0.4-0.5 and another halves are in 0.7-0.8. It shows that the general condition in the world is that the women have worse welfare compared to men. And there is a treatment gap (0.5-0.7) between the countries in two clusters.

```{r,fig.width=12}
#change -> curves
data_f = data_4 %>% group_by(Category,Year) %>% summarise(Score = mean(Yes_or_No))
data_f$Category = fct_reorder2(data_f$Category,data_f$Year,data_f$Score)
ggplot(data_f, aes(x=Year, y= Score, group = Category, color = Category)) +
  scale_y_continuous("Average Score")+
  geom_line() +
  theme_bw(base_size = 16) +
  scale_x_discrete(breaks = seq(1980,2020,5))+
  scale_color_colorblind()
```


The labels are ordered according to the average score of 30 countries in the 2020.

We can see the general trend of the women social rights in different topics from the curve plot. 
In the recent 40 years, the general trend for the women rights is increasing for every topic. And the increase started formally at around 1987-1990 for every topic. The law curve grew most rapidly from around 0.56 to 0.86, exceeding the score of job in 2020. Although the welfare score was low at the beginning, it also grew very rapidly from around 0.45 to 0.68. The growth pattern of the finance and freedom are similar, so we guess that they are somehow correlated, so do the job and marriage. Therefore, the gender discrimination problem is disappearing and it is most obvious in the welfare and law aspects.


```{r fig.width=12}
#difference -> parallel coordinates #GDP, Continents, 6 aspects.

data_GDP = data[data$Code%in%Code & data$Indicator== "NY.GDP.MKTP.CD",c(2,64)]
colnames(data_GDP) = c("Code","log_GDP")
data_GDP$log_GDP = log(data_GDP$log_GDP)
data_5 = left_join(data_3,data_GDP,by=c("Code"))
data_5 = data_5[,c(1,2,3,44,45,46,47)]
data_5 = data_5 %>% group_by(Country,Code,Continent,Category,log_GDP) %>% summarise(Score = mean(`2020`))
data_6 = spread(data_5,Category,Score)[,-c(1,2,5,6)]

parcoords(data_6,
          color = list(
            colorScale = "scaleOrdinal",
            colorBy = "Continent",
            colorScheme = c("#009E73","#F0E442", "#0072B2", "#D55E00", "#CC79A7")
          ),
          alpha = 0.5,
          withD3 = TRUE, brushMode = "1d-axes", reorderable = TRUE, rownames = FALSE)
#ColorBlind

```


Since we have more variations about the score in job, welfare, marriage and law aspects. We will only explore what potential factors influenced these variations.

Also, since some countries’ GDP are extremely high, making other countries’ GDP hard to see in the parallel coordinates, We just used the log to scale the GDP, which will not influence the absolute rank of the GDP. 

We can drag the GDP or Continent coordinates to see the correlation between the GDP or Continent and other factors we are interested in. And also we can select the range of the coordinates from top to bottom to see how the gender discrimination condition will vary when the GDP is high or the Continent is specified. 


The key findings in this part is following:

1.There is no obvious evidence that GDP will influence the gender discrimination in the social rights aspect.

2. Countries in North America will have a higher score in job while the countries in Asia will have a lower score in job. Other continents have various scores, showing no obvious trend.

3. Countries in South America and Asia will have a lower score in law while the countries in North America will have a higher score in law. Other continents have various scores, showing no obvious trend.

4. Countries in North America, Europe and Oceania will have a higher score in marriage. Other continents have various scores, showing no obvious trend.
5. Countries in Africa will have a lower score in welfare. Other continents have various scores, showing no obvious trend.

It is quite surprised that the GDP has no influence on the gender discrimination in the social rights aspect but the geographic position will influence it. Maybe the cultural reasons instead of the financial reasons will dominate the gender discrimination problem more. It quite makes sense. Although in these years we focus on solving the discrimination, it is quite hard to wipe it out since the cultural stereotypes will give us a preconception about the gender and then influence our action and decisions subconsciously.

## Education

## Employment
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 1

packages = c('tidyverse', 'dplyr', 'tidyr')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(tidyr)

#Load in data

genderStatsCountry <- read.csv("data/Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("data/Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("data/Gender_StatsData.csv")
genderStatsFootnote <- read.csv("data/Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("data/Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("data/Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV',
              'COL', 'BRA', 'CHL', 'ARG', 'URY',
              'RUS', 'ISL', 'UKR', 'BLR', 'LTU',
              'CHN', 'JPN', 'KOR', 'THA', 'MYS',
              'ZWE', 'TZA', 'NER', 'NGA', 'COG',
              'AUS', 'NZL', 'PNG', 'WSM', 'FJI')

#Processing functions
capFirst <- function(s) {
  paste(toupper(substring(s, 1, 1)), substring(s, 2), sep = "")
}

lowFirst <- function(s) {
  paste(tolower(substring(s, 1, 1)), substring(s, 2), sep = "")
}

#Employment to population ratio by sex
emp_data <- full_data %>% 
  filter(Indicator.Name.x == 'Employment to population ratio, 15+, female (%) (modeled ILO estimate)' |
           Indicator.Name.x == 'Employment to population ratio, 15+, male (%) (modeled ILO estimate)') %>% 
  select(22:24, 75:84) %>% 
  mutate_if(is.numeric, round, 1)
emp_data <- emp_data[emp_data$Country.Code %in% countries ,]
emp_data <- emp_data %>% 
  mutate(emp_data, Mean = rowMeans(select(., starts_with('X')), na.rm = TRUE)) %>% 
  select(1:3, 14) %>% 
  separate(Indicator.Name.y, c('Indicator.Name', 'Age', 'Gender'), sep = ', ') %>% 
  separate(Gender, c('Sex', 'Other'), sep = '^[^\\s]*\\K\\s') %>% 
  select(1:3, 7, 5)
emp_data$Sex <- capFirst(emp_data$Sex)

#Plot Cleveland of Employment to Population Ratio
theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(1.25)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(1.25)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 1),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#D55E00", "#0072B2", "#F0E442", "#CC79A7")

#fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),

ggplot(emp_data, aes(Mean,
                     fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),
                     color = Sex)) +
  geom_point(size = 3) +
  ggtitle("2010-2019 Average Employment Ratio for each Country by Sex") + 
  ylab("") +
  xlab("% Employed") + 
  scale_colour_manual(values = cbPalette) +
  theme_dotplot
```
The figure above for the average employment ratio by sex between 2010-2019 is ordered in descending order by male employment ratio. The purpose of this ordering is to list countries with high male employment ratios first and visualize whether or not the female employment ratios match up. Out of the 30 countries selected for the study, it is very telling that a lower female employment ratio fairly universal. The actual differences between male/female employment ratio varies drastically by country. From the figure above, it is clear that the issue is much more pronounced in countries like Samoa, Mexico, and Malaysia (among others) as opposed to countries like Canada, Congo, Rep., and Iceland. Whether or not the differences between the ratio of male/female employment varies by continent is explored in another figure below. 
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 2

#Employment in upper/middle management 
mng_data <- full_data %>% 
  filter(Indicator.Name.x == 'Female share of employment in senior and middle management (%)') %>%
  select(22:24, 75:84) %>%
  mutate_if(is.numeric, round, 1)
mng_data <- mng_data[mng_data$Country.Code %in% countries ,]
mng_data <- mng_data %>% 
  mutate(mng_data, Mean = rowMeans(select(., starts_with('X')), na.rm = TRUE)) %>% 
  select(1:3, 14) %>% 
  separate(Indicator.Name.y, c('Sex', 'Indicator.Name'), sep = ' share of ')

male_percentage <- as.data.frame(100 - mng_data$Mean)
male_percentage <- male_percentage %>% 
  rename(Mean = '100 - mng_data$Mean')
male_vals <- cbind(ï..Country.Name = mng_data$ï..Country.Name, 
                    Country.Code = mng_data$Country.Code,
                    'Sex' = 'male',
                    Indicator.Name = mng_data$Indicator.Name,
                    Mean = male_percentage)

mng_data <- rbind(mng_data, male_vals) %>% 
  select(1:2, 4:5, 3)
mng_data$Indicator.Name <- capFirst(mng_data$Indicator.Name)
mng_data$Sex <- capFirst(mng_data$Sex)
mng_data$Indicator.Name <- substr(mng_data$Indicator.Name, 1, 
                                         nchar(mng_data$Indicator.Name) - 4)
mng_data$Mean <- as.numeric(mng_data$Mean)
mng_data[is.na(mng_data)] <- 0

#Plot Cleveland of Management Data
ggplot(mng_data, aes(Mean,
                     fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),
                     color = Sex)) +
  geom_point(size = 3) +
  ggtitle("2010-2019 Average Proportion of Upper/Middle Management for each Country by Sex") + 
  ylab("") +
  xlab("Proportion") + 
  xlim(10, 90) +
  scale_colour_manual(values = cbPalette) +
  theme_dotplot

```
The figure above for the average proportion of upper/middle management by sex is ordered in descending order by male upper/middle managers. The purpose of this ordering is to list countries with the biggest discrepancy in male/female manager ratio first, which can serve as a signal for gender discrepancies in employment opportunities. The countries on the figure without data points are countries where there is no data related to this area. These missing countries were kept within the figure to signify the fact that the International Labor Organization was unable to obtain any data related to this issue from numerous countries over a 10-year span. Even without 12 of the 30 randomly selected nations from the study, there appears to be a trend that people in upper/middle management are male the vast majority of the time in almost all countries with the exception of Nicaragua (in the sample). An interesting note is that although Papua New Guinea has a near equal proportion of female employment to male employment, upper/middle managers in that country are over 75% male. On the other hand, Nicaragua has a much higher male employment ratio than female employment ratio, but most upper/middle managers are female.
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 3

packages = c('tidyverse', 'dplyr', 'grid', 'gridExtra')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)

#Load in data

genderStatsCountry <- read.csv("data/Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("data/Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("data/Gender_StatsData.csv")
genderStatsFootnote <- read.csv("data/Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("data/Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("data/Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

#Full Data of all Countries
data <- full_data %>% 
  filter(Indicator.Name.x == 'Women Business and the Law Index Score (scale 1-100)') %>% 
  select(22:23, 55:84) %>% 
  mutate_if(is.numeric, round, 1)

#NA Region Data
na_countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV')
na_data <- data[data$Country.Code %in% na_countries ,]

names(na_data) <- substring(names(na_data), 2)
na_data <- na_data %>% 
  rename(Country.Code = 'ountry.Code')

na_data <- na_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

na_data$region = 'North America'
  
#SA Region Data
sa_countries = c('COL', 'BRA', 'CHL', 'ARG', 'URY')
sa_data <- data[data$Country.Code %in% sa_countries ,]

names(sa_data) <- substring(names(sa_data), 2)
sa_data <- sa_data %>% 
  rename(Country.Code = 'ountry.Code')

sa_data <- sa_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

sa_data$region = 'South America'

#EU Region Data
eu_countries = c('RUS', 'ISL', 'UKR', 'BLR', 'LTU')
eu_data <- data[data$Country.Code %in% eu_countries ,]

names(eu_data) <- substring(names(eu_data), 2)
eu_data <- eu_data %>% 
  rename(Country.Code = 'ountry.Code')

eu_data <- eu_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

eu_data$region = 'Europe'

#AS Region Data 
as_countries = c('CHN', 'JPN', 'KOR', 'THA', 'MYS')
as_data <- data[data$Country.Code %in% as_countries ,]

names(as_data) <- substring(names(as_data), 2)
as_data <- as_data %>% 
  rename(Country.Code = 'ountry.Code')

as_data <- as_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

as_data$region = 'Asia'

#AF Region Data 
af_countries = c('ZWE', 'TZA', 'NER', 'NGA', 'COG')
af_data <- data[data$Country.Code %in% af_countries ,]

names(af_data) <- substring(names(af_data), 2)
af_data <- af_data %>% 
  rename(Country.Code = 'ountry.Code')

af_data <- af_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

af_data$region = 'Africa'

#OC Region Data
oc_countries = c('AUS', 'NZL', 'PNG', 'WSM', 'FJI')
oc_data <- data[data$Country.Code %in% oc_countries ,]

names(oc_data) <- substring(names(oc_data), 2)
oc_data <- oc_data %>% 
  rename(Country.Code = 'ountry.Code')

oc_data <- oc_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

oc_data$region = 'Oceania'


#Plotting
full_data <- rbind(na_data, sa_data, eu_data, as_data, af_data, oc_data)

df <- split(full_data, f = full_data$region)

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#D55E00", "#0072B2", "#F0E442", "#CC79A7")

na <- ggplot(data = df$'North America', 
             aes(x = year, y = index_score, color = ..Country.Name, group = Country.Code)) + 
  geom_point() +
  geom_line() +
  facet_wrap(~region, ncol = 1) +
  labs(x = 'Year', y = 'Index Score') +
  theme_grey(16) +
  ylim(30, 100) +
  scale_colour_manual(values = cbPalette) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1),
        legend.title = element_blank())

sa <- na %+% df$'South America'
eu <- na %+% df$'Europe'
as <- na %+% df$'Asia'
af <- na %+% df$'Africa'
oc <- na %+% df$'Oceania'

grid.arrange(na,
             sa,
             eu,
             as,
             af,
             oc,
             top = textGrob("Women Business and the Law Index Score by Region Over the Years",
                            gp=gpar(fontsize=20,font=3)))

```
The Women, Business and the Law Score is an index that analyzes laws and regulations that affect women’s economic inclusion in 190 economies. It is composed of eight indicators revolving around women’s interactions with the law as they progress through their careers and make economic decisions. The eight indicators are: mobility, workplace, pay, marriage, parenthood, entrepreneurship, assets, and pension. The index ranges from 1-100 with 100 being the highest score, or equal legal rights with men in measured areas.  
  
The figure above displays the Women, Business and the Law index scores over a 30-year period faceted by continent with 5 countries per continent. The figure indicates a universal trend upwards in the index score over nearly all countries in every continent, although the degree of improvement varies drastically. For example, countries like the Congo, Rep. saw no improvement since 1900 and Papua New Guinea saw very marginal improvements, whereas countries like Brazil and Lithuania saw drastic improvements over the same time frame. It is very uncommon for the index score to go downwards once it has gone up, although it has happened on occasions like in 2004 for Lithuania, 2014 in Ukraine, and 2008 in China. A significant marker for gender inequality in employment is that only 2 of the 30 selected countries ever achieved a score of 100, signifying equal legal rights with men in the measured areas of the index. Although women’s legal rights in the eight indicator areas have improved worldwide since 1990, there are very few countries where equality is achieved as of now. 

```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 4

packages = c('tidyverse', 'dplyr', 'grid', 'gridExtra', 'scales', 'patchwork')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)
library(scales)
library(patchwork)

#Load in data


data = read.csv("data/Gender_StatsData.csv",header = TRUE,encoding="UTF-8")
colnames(data)=c("Country","Code","Indicator_details","Indicator",as.character(1960:2020))

Code = c("CAN","USA","MEX","NIC","SLV","COL","BRA","CHL","ARG","URY","RUS","ISL","UKR","BLR","LTU","CHN","JPN","KOR","THA","MYS","COG","ZWE","TZA","NER","NGA",'AUS','NZL','PNG','WSM','FJI')
Continent = rep(c("North America","South America","Europe","Asia","Africa","Oceania"),rep(5,6))
data_continent = data.frame(cbind(Code,Continent))

data_GDP = data[data$Code%in%Code & data$Indicator== "NY.GDP.PCAP.CD",c(2,64)]
colnames(data_GDP) = c("Code","GDP")
data_GDP$GDP = log(data_GDP$GDP)


data_incom = read.csv("data/Gender_StatsCountry.csv",header = TRUE,encoding="UTF-8")[,c(1,9)]
colnames(data_incom) = c("Code","Income")
data_INC = data_incom[data_incom$Code%in%Code,]
data_EMP = data[data$Code%in%Code & data$Indicator == "SL.TLF.CACT.FM.ZS",c(2,55:64)]
colnames(data_EMP) = c("Code",as.character(c(2010:2019)))
data_EMP = data_EMP%>%pivot_longer(cols = starts_with("1") | starts_with("2"),names_to = "Year",values_to = "ratio")
data_EMP = data_EMP%>% group_by(Code) %>% summarise(Mean = sum(ratio)/n())
data_EMP_AN = left_join(left_join(left_join(data_EMP,data_GDP,by=c("Code")),data_continent,by=c("Code")),data_INC,by = c("Code"))

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#0072B2", "#CC79A7")

ggplot(data_EMP_AN,aes(y=Mean,x=GDP, color = Continent))+
  geom_point(size = 3) +
  ggtitle('2010-2019 Average Female/Male Labor Force Participation Rate vs. Log-Scale GDP') +
  scale_colour_manual(values = cbPalette) +
  scale_x_continuous("Log(GDP)") +
  scale_y_continuous("Average of Female/Male Labor Force Participation Rate") +
  theme_grey(16) +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

```
The figure above displays the ratio of female to male employment ratio vs. GDP on a log scale for each country color coded by continent. There appears to no correlation overall between a country’s female/male labor force participation rate to the country’s GDP as the data points are scattered throughout as opposed to group along a linear line. However, when looking at individual continents separately, female/male labor force participation rate does seem correlated with GDP in continents like Europe, Africa, and North America (trend is not really observable for Asia, Oceania, and South America). This suggests that the correlation between female/male employment ratio may have a lot to do with geographical location of the country, and depending on the location, the ratio may be correlated with GDP.