--- 
title: "Analyzing Gender Discrepancies Over Time Using Social, Educational, and Employment Trends"
author: "Vincent Chang, Xiaoyan Li, Yaojia Ye"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

In 2016, the American presidential election became the most heated topic around the world. On one hand, the support rates between Donald Trump and Hillary Clinton were very close, making the result hard to predict. If Hillary won the election, she would be the first women president in the American history. Hillary did not end up winning the presidential election but it is noteworthy that an extremely imbalanced gender distribution existed in specific working fields such as the president. Whether the gender discrimination exists worldwide and to what extent is gender discrimination present seem to be a interesting subject worthy to be explored. In this project, we want to gain a overall perception on gender discrimination in the world by mainly seeking the answers for three sub-questions through exploratory data analysis:

1) How do women's social status compare to men's social status in modern society from a variety of standpoints like social rights, education, and employment?

2) How has women's social status changed in the last 30 years (years data is available for) regarding social rights, education, and employment?  

3) What are the differences and developments in women's social status between different countries in recent years? Does GDP (or other developmental / geographical factors) have an influence on equal rights. What are plausible explanations for gender discrepancies from the perspective of different countries?  

Owing to the data availability for countries, we used only 30 countries with high quality data to do the analysis, each country was randomly chosen from 6 continents. We hope the situations in these countries can somehow reflect the gender discrepancies as a whole around the world.

<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources

The data for this project is sourced from the World Bank. The World Bank is a group dedicated to providing high-quality statistical data on numerous topics globally to aid development and policy making. A lot of the information within the World Bank’s databases come directly from the statistical systems of member countries, which means the accuracy of the information is heavily dependent on how well these national systems perform their data collection. As a result, a high priority of the World Bank is to help improve statistical infrastructures of developing countries by preparing national strategies for the development of statistics “as recommended in the Marrakech Action Plan for Statistics”. Data collection is currently a joint effort between the World Bank and many other international statistical communities including the United Nations (UN), the Organization for Economic Co-Operation and Development (OECD), the International Monetary Fund (IMF), and regional development banks. Transnational data is also collected by the World Bank through globally coordinated programs like the Marrakech Action Plan for Statistics and PARIS21.  

For all the data collected, the World Bank ensures data consistency and reliability through:  
  -	Participation in the UN Statistical Commission and other statistical forums to develop appropriate frameworks, guidance, and standards of good practice for statistics  
  -	Building consensus and defining internationally agreed indicators such as the indicators for the Millennium Development Goals  
  -	Establishing data exchange processes and methods  
  -	Assembling, analyzing, and disseminating data online and in print  

The gender statistics data used for this project has 164,112 rows and 66 columns. Each entry represents a single indicator for a country. In total, there are 263 countries and each has 624 different indicators. Each indicator is a record item, eg. GDP, and have numeric data recorded from 1960 to 2020 (with more missing values in older years). In each row, it has 4 string variables: country name, country code, a description of indicator, and an indicator code. Some indicators have a boolean form like the yes or no statement, with the numeric data entries 1 as yes, 0 as no. The indicators can be categorized into 5 types.  
  
1) General information about the country, like GDP, fertility rate, poverty, population  
2) Education related items, like school enrollment  
3) Employment and unemployment  
4) Specific female social status indicator, like if a women can apply a passport in a same way as a man  
5) Health related indicator, like life expectancy and rate of overweight  

One problem in this data file is the high proportion of missing values. It is understandable that some country may not have records on some specific indicators, or the records are missing for some years. If we used the data from all countries, we may have too many missing values that makes our results unreliable. Thus, we decided to take a sample of countries that has relatively fewer missing values on the indicators we aimed to use.

*Info comes from: https://data.worldbank.org/about

<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

Data for this project is sourced from the following link:   https://datacatalog.worldbank.org/dataset/gender-statistics  
  
The CSV files were downloaded from the "Data & Resources" tab and imported directly into RStudio for processing. The specific steps taken to process the data for visualization differed between the three topics studied for this project: social rights, education, and employment. Since there is information for over 200 regions, the project selected 30 to analyze their changes over time. Out of the 30 selected countries, 5 belonged to each continent. These countries were randomly selected.  
  
The selected countries are:  
North America: Canada, El Salvador, Mexico, Nicaragua, United States  
South America: Argentina, Brazil, Chile, Colombia, Uruguay  
Europe: Belarus, Iceland, Lithuania, Russian Federation, Ukraine  
Asia: China, Japan, Korea, Malaysia, Thailand  
Africa: Rep. Congo, Niger, Nigeria, Tanzania, Zimbabwe  
Oceania: Australia, Fiji, New Zealand, Papua New Guinea, Samoa  
  
For social rights, the dataset has a lot of questions related to the women social rights with binary labels attached. We divided 24 relevant rights into 6 topics: 2 for Finance, 3 for Freedom, 8 for Law, 3 for Marriage, 4 for Job and 4 for Welfare. If the answer to the question is yes, then the label is 1 (otherwise the label is 0). We averaged the binary numbers in each topic to get the averaged score for each topic. We then used the distribution of the scores for individual countries  to analyze social rights issues as presented in the introduction (details are in the check_list.csv).
  
For education, we selected indicators that represented the female and male education status across the time like school enrollment and rate of graduation. Moreover, in order to detect the correlation between female education status and other factors, we also selected GDP, fertility rate,  female rural population and government expenditure on education department. To reduce the effect of missing values, we use average scores across multiple years instead of using data from specific year on most of the plot except for the time-series plots.
  
For employment, the specific features selected from the data set included the employment to population ratio of males and females aged 15+ from model ILO estimates, the female share of employment in upper/middle management, the Women business and the law index, and the female/male employment ratio. The Women business and the law index had no missing values. 2010-2019 averages were used for data visualization for the other features where some values were missing. 

<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values

## Social Rights  
There is no missing data for the social rights aspect analysis.

## Education 
```{r, include=FALSE}
SOLUTIONS <- TRUE
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, 
                      eval = SOLUTIONS, cache = FALSE)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(forcats)
library(parcoords)
library(reshape2)
library(GGally)
library(mi)
```

```{r echo=FALSE}
file <- "data/Gender_StatsData.csv"
data <- read.csv(file)
```

```{r}
country <- data %>% 
  filter(Country.Code == "CAN" | Country.Code == "USA" | Country.Code == "MEX" | Country.Code == "NIC" | Country.Code == "SLV" | Country.Code == "COL" | Country.Code == "BRA" | Country.Code == "CHL" | Country.Code == "ARG" | Country.Code == "URY" | Country.Code == "RUS" | Country.Code == "ISL" | Country.Code == "UKR" | Country.Code == "BLR" | Country.Code == "LTU" | Country.Code == "CHN" | Country.Code == "JPN" | Country.Code == "KOR" | Country.Code == "THA" | Country.Code == "MYS" | Country.Code == "COG" | Country.Code == "ZWE" | Country.Code == "TZA" | Country.Code == "NER" | Country.Code == "NGA" | Country.Code == "AUS" | Country.Code == "NZL" | Country.Code == "PNG" | Country.Code == "WSM" | Country.Code == "FJI" ) 
```

```{r}
data_q3 <- country %>% 
  filter(Indicator.Code == "SP.DYN.TFRT.IN" | Indicator.Code == "NY.GDP.MKTP.CD" |  Indicator.Code == "SE.SCH.LIFE.FE" |  Indicator.Code == "SE.XPD.TOTL.GD.ZS" |  Indicator.Code == "SP.RUR.TOTL.FE.ZS")
```

```{r}
data_q3 <- data_q3 %>% mutate_if(is.factor, funs(case_when( 
    . == "SP.DYN.TFRT.IN" ~ "FertilityRate",
    . == "NY.GDP.MKTP.CD" ~ "GDP(log in billion)",
    . == "SE.SCH.LIFE.FE" ~ "ExpSchoolYear",
    . == "SE.XPD.TOTL.GD.ZS" ~ "EducationSpend",
    . == "SP.RUR.TOTL.FE.ZS" ~ "RuralPop",
    . == "CAN" ~ "North America", 
    . == "USA" ~ "North America",
    . == "MEX" ~ "North America",
    . == "NIC" ~ "North America",
    . == "SLV" ~ "North America",
    . == "COL" ~ "South America",
    . == "BRA" ~ "South America",
    . == "CHL" ~ "South America",
    . == "ARG" ~ "South America",
    . == "URY" ~ "South America",
    . == "RUS" ~ "Europe",
    . == "ISL" ~ "Europe",
    . == "UKR" ~ "Europe",
    . == "BLR" ~ "Europe",
    . == "LTU" ~ "Europe",
    . == "CHN" ~ "Asia",
    . == "JPN" ~ "Asia",
    . == "KOR" ~ "Asia",
    . == "THA" ~ "Asia",
    . == "MYS" ~ "Asia",
    . == "COG" ~ "Africa",
    . == "ZWE" ~ "Africa",
    . == "TZA" ~ "Africa",
    . == "NER" ~ "Africa",
    . == "NGA" ~ "Africa",
    . == "AUS" ~ "Oceania",
    . == "NZL" ~ "Oceania",
    . == "PNG" ~ "Oceania",
    . == "WSM" ~ "Oceania",
    . == "FJI" ~ "Oceania",
    T ~ as.character(.))))
```


```{r}
df <- data_q3 %>% 
  select(ï..Country.Name,Country.Code, Indicator.Code, c(40:64)) %>%
  group_by(ï..Country.Name, Country.Code, Indicator.Code)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
```


```{r}
value <- rowMeans(df[4:28],  na.rm = TRUE)
df$value <- value
```

```{r}
df <- dcast(df, ï..Country.Name + Country.Code ~ Indicator.Code)
colnames(df)[colnames(df) == "Country.Code"] <- "Continent"
```

```{r}
df <- df[ , c(1, 2, 6, 5, 3, 7, 4)] 
df[13,7] <- NA
df[21,5] <- NA
```

```{r, echo = FALSE, fig.height = 10, fig.width = 10}
library(mi)
x <- missing_data.frame(df)
image(x)
```


The above heatmap shows the missing value for the parallel coordinates in part 5.2. There are only two missing values in this data frame that has 30 rows and 7 columns. As all the data entries (except for country name and continent) are average scores across 25 years, the missing data here actually represents a 25 years data missing in our original data file. Thus, this heatmap implies that in the 30 countries we chose, there is one country have no data entry for education spend and one country has no data for expected school year in recent 25 years.

## Employment 
```{r, echo = FALSE, fig.height = 10, fig.width = 10}
#Vincent Figure 1
defaultW <- getOption("warn") 
options(warn = -1) 

packages = c('tidyverse', 'dplyr', 'tidyr', 'mi', 'stringr')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(tidyr)
library(mi)
library(stringr)

#Load in data

genderStatsCountry <- read.csv("data/Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("data/Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("data/Gender_StatsData.csv")
genderStatsFootnote <- read.csv("data/Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("data/Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("data/Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV',
              'COL', 'BRA', 'CHL', 'ARG', 'URY',
              'RUS', 'ISL', 'UKR', 'BLR', 'LTU',
              'CHN', 'JPN', 'KOR', 'THA', 'MYS',
              'ZWE', 'TZA', 'NER', 'NGA', 'COG',
              'AUS', 'NZL', 'PNG', 'WSM', 'FJI')

emp_data <- full_data %>% 
  filter(Indicator.Name.x == 'Employment to population ratio, 15+, female (%) (modeled ILO estimate)') %>% 
  select(22:24, 75:84) %>% 
  mutate_if(is.numeric, round, 1)
emp_data <- emp_data[emp_data$Country.Code %in% countries ,]
emp_data <- emp_data %>% 
  separate(Indicator.Name.y, c('Indicator.Name', 'Age', 'Gender'), sep = ', ') %>% 
  separate(Gender, c('Sex', 'Other'), sep = '^[^\\s]*\\K\\s')

emp_data_m <- full_data %>% 
  filter(Indicator.Name.x == 'Employment to population ratio, 15+, male (%) (modeled ILO estimate)') %>% 
  select(22:24, 75:84) %>% 
  mutate_if(is.numeric, round, 1)
emp_data_m <- emp_data_m[emp_data_m$Country.Code %in% countries ,]
emp_data_m <- emp_data_m %>% 
  separate(Indicator.Name.y, c('Indicator.Name', 'Age', 'Gender'), sep = ', ') %>% 
  separate(Gender, c('Sex', 'Other'), sep = '^[^\\s]*\\K\\s')

mng_data <- full_data %>% 
  filter(Indicator.Name.x == 'Female share of employment in senior and middle management (%)') %>%
  select(22:24, 75:84) %>%
  mutate_if(is.numeric, round, 1)
mng_data <- mng_data[mng_data$Country.Code %in% countries ,]

lab_data <- full_data %>% 
  filter(Indicator.Name.x == 'Ratio of female to male labor force participation rate (%) (modeled ILO estimate)') %>%
  select(22:24, 75:84) %>%
  mutate_if(is.numeric, round, 1)
lab_data <- lab_data[lab_data$Country.Code %in% countries ,]

df <- emp_data %>% 
  select(1, 7:16)
names(df)[substr(names(df), 1, 1) == 'X'] <- paste(substr(names(df), 1, 5), 'var1')

df2 <- emp_data_m %>% 
  select(1, 7:16)
names(df2)[substr(names(df2), 1, 1) == 'X'] <- paste(substr(names(df2), 1, 5), 'var2')

df3 <- mng_data %>% 
  select(1, 4:13)
names(df3)[substr(names(df3), 1, 1) == 'X'] <- paste(substr(names(df3), 1, 5), 'var3')

df4 <- lab_data %>% 
  select(1, 4:13)
names(df4)[substr(names(df4), 1, 1) == 'X'] <- paste(substr(names(df4), 1, 5), 'var4')

df_final <- merge(df, df2, by.x = 'ï..Country.Name', by.y = 'ï..Country.Name')
df_final <- merge(df_final, df3, by.x = 'ï..Country.Name', by.y = 'ï..Country.Name')
df_final <- merge(df_final, df4, by.x = 'ï..Country.Name', by.y = 'ï..Country.Name')

x <- missing_data.frame(df_final)
fig <- image(x)
fig

options(warn = defaultW)
```
For the data used to analyze employment trends, v1 signifies the employment to population ratio of females 15+, v2 signifies the employment to population ratio of males 15+, v3 signifies the proportion of females in upper/middle management, and v4 signifies the Women business and the law index scores. There are no missing values for v1, v2, and v4 as they are model ILO estimates, so some values are interpolated by the World Bank. Even though v3 has a ton of missing values, the proportion of females to males in upper/middle management relative to the proportion of females to males in the workforce appears to be a strong indicator for discrepancies in employment opportunities, so a decision was made to keep the feature for analysis. Out of the 30 countries selected, 12 countries had no information on the proportion of females in upper/middle management and they were excluded from the analysis that followed.

<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

By exploring the gender data, we want to analyze the gender discrepancies to answer the following questions:

1) How do women's social status compare to men's social status in modern society from a variety of standpoints like social rights, education, and employment?

2) How has women's social status changed in the last 30 years (years data is available for) regarding social rights, education, and employment?  

3) What are the differences and developments in women's social status between different countries in recent years? Does GDP (or other developmental / geographical factors) have an influence on equal rights. What are plausible explanations for gender discrepancies from the perspective of different countries?  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE,message = FALSE)
```

## Social

To discuss the social rights for women in the world these years.

To explore the details about the social rights, the heatmap is a good choice. The y axis is ordered by the average scores of countries from high to low. The x axis is ordered by the average scores of the topics.

```{r ,fig.width=12, fig.height = 12}
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(parcoords)

data = read.csv("data/Gender_StatsData.csv",header = TRUE,encoding="UTF-8")
colnames(data)=c("Country","Code","Indicator_details","Indicator",as.character(1960:2020))

Code = c("CAN","USA","MEX","NIC","SLV","COL","BRA","CHL","ARG","URY","RUS","ISL","UKR","BLR","LTU","CHN","JPN","KOR","THA","MYS","COG","ZWE","TZA","NER","NGA",'AUS','NZL','PNG','WSM','FJI')
Continent = rep(c("North_America","South_America","Europe","Asia","Africa","Oceania"),rep(5,6))
data_continent = data.frame(cbind(Code,Continent))

Indicator = c("SG.APL.PSPT.EQ","SG.HLD.HEAD.EQ","SG.LOC.LIVE.EQ","SG.GET.JOBS.EQ","SG.OBT.DVRC.EQ","SG.OPN.BANK.EQ","SG.BUS.REGT.EQ","SG.CNT.SIGN.EQ","SG.HME.TRVL.EQ","SG.CTR.TRVL.EQ","SG.NGT.WORK.EQ","SG.DNG.WORK.DN.EQ","SG.IND.WORK.EQ","SG.REM.RIGT.EQ","SG.AGE.FUPN.EQ","SG.AGE.PAPN.EQ","SH.MMR.LEVE.GT","SG.LAW.ASST.AR","SG.LAW.CRDD.GR","SG.AGE.MRET.EQ","SG.ABS.PENB","SG.LEG.SXHR.EM","SG.LEG.DVAW","SG.LAW.OBHB.MR.NO")
Category = c("law","marriage","freedom","job","marriage","finance","finance","law","freedom","freedom","job","job","job","marriage","welfare","welfare","welfare","law","law","law","welfare","law","law","law")
data_cat = data.frame(cbind(Indicator,Category))

data_1 = data[data$Code%in%Code & data$Indicator%in%Indicator,c(1,2,4,25:65)]
data_2 = left_join(data_1,data_continent,by=c("Code"))
data_3 = left_join(data_2,data_cat,by=c("Indicator"))

data_4 = data_3 %>%pivot_longer(cols = starts_with("1") | starts_with("2"),names_to = "Year",values_to = "Yes_or_No")
data_c = data_4 %>% group_by(Country,Code,Continent,Category,Year) %>% summarise(Score = mean(Yes_or_No))

#data_e for reorder
data_tot_s_y = data_c[data_c$Year == 2020,] %>% group_by(Country) %>% summarise(Tot_S_Y = sum(Score))
data_tot_s_x = data_c[data_c$Year == 2020,] %>% group_by(Category) %>% summarise(Tot_S_X = sum(Score))
data_e = left_join(data_c[data_c$Year == 2020,],data_tot_s_y,by=c("Country"))
data_e = left_join(data_e,data_tot_s_x,by = c("Category"))

ggplot(data_e,aes(y=fct_reorder(Country,Tot_S_Y),x=fct_rev(fct_reorder(Category,Tot_S_X))))+
  geom_tile(aes(fill=Score)) +
  scale_fill_distiller(palette='Reds',direction = 1) +
  scale_x_discrete("Score")+
  scale_y_discrete("Country")+
  theme_classic(base_size = 20) +
  ggtitle("The Social Right Score in 30 countries") +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank())
```

We can see that Iceland, Canada and New Zealand actually both achieved a score of 1 in 6 topics, meaning that there are no gender discrimination in the social rights aspect in these three countries. Compared to that, Niger has no score of 1 in every topic, meaning there is still some significant gender discrepancies there, especially for marriage where the score is close to 0. This means men have more power to dominate the marriage in Niger. Another point worth paying attention to is that jobs also have a very low score in Ukraine while other scores are quite high, meaning that women face more significant gender discrimination in the workplace as opposed to other social aspects there.

Moreover, there are some countries have really similar social rights conditions for women. For example, Canada and Iceland and New Zealand, United States and Mexico, Nicaragua and Lithuania, Zimbabwe and El Salvador, Fiji and Colombia, China and Belarus, Russian Federation and Papua New Guinea. We will discuss what kinds of factor contributes to the distribution of the scores like this.

Generally, women tend to have the same rights as the men in the finance aspect since only 1 country did not achieve 1 in finance. For the freedom part, only 4 countries did not achieve 1, meaning that the majority of the countries do not keep women from doing something which men can do. However, in terms of the social welfare, the women tend to have a worse welfare guaranteed compared to men in these years with only 5 countries achieving score 1. As for law, marriage and job, different countries have a wide variety of different conditions.

```{r fig.width=12}

ggplot(data_e, aes(Score)) +
  geom_histogram(color = "grey", 
                 fill = "lightBlue",
                 binwidth = 0.1, 
                 center = 0.05, 
                 closed = "right")+
  facet_wrap(~ Category)+
  ggtitle("The Distribution of Social Right Score") +
  scale_x_continuous(breaks=seq(0,1,0.1))+
  theme_bw(base_size = 20) 

#situation -> heatmap #6 aspects vs 25 countries, scores
```

Since the heatmap do not show the detailed score distribution of topics clearly. We can explore the general distribution more carefully by looking at the bar charts. We can see that for the finance part ad freedom part, the majorities of countries has a score of 1. For the job aspect, we can see that except for the score 1 and the outlier laying between 0.2 to 0.3, there are two clusters of countries distributed in the range of 0.4-0.5 and 0.7-0.8, meaning there is a gap (0.5-0.7) between the countries for the women social status in the job aspect. For law, we can see the range of the clusters are nested in the right corner except the outlier laying between 0.3 to 0.4, meaning the general scores are high in the law aspect and the difference of the condition are small between the countries. For marriage, only 3 countries are outliers, and the majority countries are landed in score 1 with approximately 7 countries having 0.6-0.7 scores. For welfare, we can see there are mainly 3 clusters, for 0.4-0.5, 0.7-0.8, 0.9-1.0 respectively. And nearly halves are in 0.4-0.5 and another halves are in 0.7-0.8. It shows that the general condition in the world is that the women have worse welfare compared to men. And there is a treatment gap (0.5-0.7) between the countries in two clusters.

```{r,fig.width=12}
#change -> curves
data_f = data_4 %>% group_by(Category,Year) %>% summarise(Score = mean(Yes_or_No))
data_f$Category = fct_reorder2(data_f$Category,data_f$Year,data_f$Score)
ggplot(data_f, aes(x=Year, y= Score, group = Category, color = Category)) +
  scale_y_continuous("Average Score")+
  geom_line() +
  ggtitle("The Social Right Score in last 40 years") +
  theme_bw(base_size = 20) +
  scale_x_discrete(breaks = seq(1980,2020,5))+
  scale_color_colorblind()
```


The labels are ordered according to the average score of 30 countries in the 2020.

We can see the general trend of the women social rights in different topics from the curve plot. 
In the recent 40 years, the general trend for the women rights is increasing for every topic. And the increase started formally at around 1987-1990 for every topic. The law curve grew most rapidly from around 0.56 to 0.86, exceeding the score of job in 2020. Although the welfare score was low at the beginning, it also grew very rapidly from around 0.45 to 0.68. The growth pattern of the finance and freedom are similar, so we hypothesize that they are somehow correlated. Job and marriage seems to follow a similar trend to finance and freedom. Therefore, the gender discrimination problem is disappearing and it is most obvious in the welfare and law aspects.


```{r fig.width=12}
#difference -> parallel coordinates #GDP, Continents, 6 aspects.

data_GDP = data[data$Code%in%Code & data$Indicator== "NY.GDP.MKTP.CD",c(2,64)]
colnames(data_GDP) = c("Code","log_GDP")
data_GDP$log_GDP = log(data_GDP$log_GDP)
data_5 = left_join(data_3,data_GDP,by=c("Code"))
data_5 = data_5[,c(1,2,3,44,45,46,47)]
data_5 = data_5 %>% group_by(Country,Code,Continent,Category,log_GDP) %>% summarise(Score = mean(`2020`))
data_6 = spread(data_5,Category,Score)[,-c(1,2,5,6)]

parcoords(data_6,
          color = list(
            colorScale = "scaleOrdinal",
            colorBy = "Continent",
            colorScheme = c("#009E73","#F0E442", "#0072B2", "#D55E00", "#CC79A7")
          ),
          alpha = 0.5,
          withD3 = TRUE, brushMode = "1d-axes", reorderable = TRUE, rownames = FALSE)
#ColorBlind

```

Since we have more variations in score for job, welfare, marriage and law aspects. We will only explore what potential factors influenced these variations.

Also, since some countries’ GDP are extremely high, making other countries’ GDP hard to see in the parallel coordinates, We used the log to scale the GDP, which will not influence the absolute rank of the GDP. 

We can drag the GDP or Continent coordinates to see the correlation between the GDP or Continent and other factors we are interested in. And also we can select the range of the coordinates from top to bottom to see how the gender discrimination condition will vary when the GDP is high or the Continent is specified. 

The key findings in this part is following:

1.There is no obvious evidence that GDP will influence the gender discrimination in the social rights aspect.

2. Countries in North America will have a higher score in job while the countries in Asia will have a lower score in job. Other continents have various scores, showing no obvious trend.

3. Countries in South America and Asia will have a lower score in law while the countries in North America will have a higher score in law. Other continents have various scores, showing no obvious trend.

4. Countries in North America, Europe and Oceania will have a higher score in marriage. Other continents have various scores, showing no obvious trend.

5. Countries in Africa will have a lower score in welfare. Other continents have various scores, showing no obvious trend.

It is quite surprised that the GDP has no influence on the gender discrimination in the social rights aspect but the geographic position will influence it. Maybe the cultural reasons instead of the financial reasons will dominate the gender discrimination problem more. It quite makes sense. Although in these years we focus on solving the discrimination, it is quite hard to wipe it out since the cultural stereotypes will give us a preconception about the gender and then influence our action and decisions subconsciously.


## Education

```{r echo=FALSE}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(forcats)
library(parcoords)
library(reshape2)
library(GGally)
library(mi)
file <- "data/Gender_StatsData.csv"
data <- read.csv(file)
```

```{r}
country <- data %>% 
  filter(Country.Code == "CAN" | Country.Code == "USA" | Country.Code == "MEX" | Country.Code == "NIC" | Country.Code == "SLV" | Country.Code == "COL" | Country.Code == "BRA" | Country.Code == "CHL" | Country.Code == "ARG" | Country.Code == "URY" | Country.Code == "RUS" | Country.Code == "ISL" | Country.Code == "UKR" | Country.Code == "BLR" | Country.Code == "LTU" | Country.Code == "CHN" | Country.Code == "JPN" | Country.Code == "KOR" | Country.Code == "THA" | Country.Code == "MYS" | Country.Code == "COG" | Country.Code == "ZWE" | Country.Code == "TZA" | Country.Code == "NER" | Country.Code == "NGA" | Country.Code == "AUS" | Country.Code == "NZL" | Country.Code == "PNG" | Country.Code == "WSM" | Country.Code == "FJI" ) 
```


```{r}
data_q1 <- country %>% 
  filter(Indicator.Code == "SE.PRM.CUAT.FE.ZS" | Indicator.Code == "SE.PRM.CUAT.MA.ZS" | Indicator.Code == "SE.SEC.CUAT.LO.FE.ZS" | Indicator.Code == "SE.SEC.CUAT.LO.MA.ZS" |  Indicator.Code == "SE.SEC.CUAT.PO.FE.ZS" | Indicator.Code == "SE.SEC.CUAT.PO.MA.ZS") %>% 
  select(c(3,4,56:65))
```


```{r}
df <- data_q1 %>% 
  group_by(Indicator.Code)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
proportion <- rowMeans(df[2:11],  na.rm = TRUE)
Gender <- c("Female", "Male", "Female", "Male", "Female", "Male")
educ <- c("Primary", "Primary", "Lower secondary", "Lower secondary", "Post secondary", "Post secondary")
df_1 <- data.frame(Gender, educ, proportion)
```


```{r, fig.width = 12, fig.height = 12}
order <- c("Primary", "Lower secondary", "Post secondary")
ggplot(df_1, aes(factor(educ, levels = order), proportion, fill = Gender)) + 
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_brewer(palette = "Set1") +
  xlab("Education level") +  
  ylab("Proportion in each population") +
  ggtitle("Education level completed in population of 25+") +
  theme_grey(16)
```

First, we want to focus on the difference between female and male in education status, thus using a bar plot is a good choice as it clearly shows the differences between these two groups in each category. This bar plot shows the proportion in female and male population over 25 years old based on education levels they at least complete. We divided the education levels into three categories: primary(grade 1-6), lower secondary(grade 7-12) and post secondary(all education after high school). In order to reduce the influences of missing values, We took the average of 30 countries in our sample during the recent 10 years. Thus this plot shows the overall education status in female and male population instead of focusing on a specific location or a specific year. 

By observing the plot, the proportion decreases for both female and male population as the education level becomes higher. Over 80 percent female and male completed at least a primary level education, and over 60 percent of each population completed at least the lower secondary, while only less than 30 percent completed the post secondary education. This observation is consistent with our common sense: there will be less proportion of people in higher education. 

Moreover, the data of female and male groups tend to have small difference in all three education levels. In the first two categories, primary and lower education, male population showed higher proportion of people completing this education level than female population, whereas in the last category, the plot showed there are higher proportion of female who graduated from a post secondary school than males. If we look at the scale of difference in the first two group, the difference in the lower secondary is smaller than the difference in primary. Thus, based on these two observations, we find that males tend to have more opportunities to take a basic level education than females. Also, among the population that completed the previous level of education, relative more females complete the next level education than males.



```{r}
data_q2 <- country %>% 
  filter(Indicator.Code == "SE.TER.GRAD.FE.SI.ZS" | Indicator.Code == "UIS.FGP.5T8.F400" | Indicator.Code == "SE.TER.GRAD.FE.ED.ZS" | Indicator.Code == "SE.TER.GRAD.FE.SV.ZS" |  Indicator.Code == "SE.TER.GRAD.FE.AG.ZS" | Indicator.Code == "SE.TER.GRAD.FE.HU.ZS") %>%
  select(c(3,4,43:63))
```


```{r}
data_q2 <- data_q2 %>% 
  group_by(Indicator.Code)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
```


```{r}
data_q2 <- data_q2 %>% mutate_if(is.factor, funs(case_when( 
    . == "SE.TER.GRAD.FE.SI.ZS" ~ "STEM",
    . == "UIS.FGP.5T8.F400" ~ "Business, Administration and Law",
    . == "SE.TER.GRAD.FE.ED.ZS" ~ "Education",
    . == "SE.TER.GRAD.FE.SV.ZS" ~ "Services",
    . == "SE.TER.GRAD.FE.HU.ZS" ~ "Arts and Humanities",
    . == "SE.TER.GRAD.FE.AG.ZS" ~ "Agriculture, Forestry and Fisheries", T ~ as.character(.))))
```

```{r}
colnames(data_q2)<-gsub("X","",colnames(data_q2))
```

```{r}
library(reshape2)
data_q2 <- melt(data_q2)
```

```{r}
data_q2$Indicator.Code = fct_reorder2(data_q2$Indicator.Code, data_q2$variable, data_q2$value)
```

```{r fig.width=12}
ggplot(data_q2, aes(variable, value, group=factor(Indicator.Code), color=factor(Indicator.Code))) +  geom_line() +
  ggtitle("Female share of graduates from different study area") +
  labs(x = "Year", y = "Percent") + 
  theme_grey(16) +
  scale_x_discrete(breaks = seq(2000, 2019, 3)) +
  theme(legend.title = element_blank()) + 
  scale_color_colorblind()
```


This line plot shows us how the percentage of female graduates in different study areas changed over the 20 years (1998 - 2018). The reason to not include the year 2019 and 2020 is because of the high number of missing values in these two years. As we can see, there are big differences between different majors. For example, there are always more than 70 percent of female graduates from Education and only about 30 percent of female graduates from STEM majors. The different study area also shows various trends. We see slightly decreasing trends in Education, Art & Humanities and STEM majors, and overall increasing in Business & Law and Agriculture & Forest. The percentage of female graduates in Services increased a lot before 2005 and later decreased in recent years, but overall, it still shows an increasing trend. Based on the plot, there seems to still have some gender differences among the various subject areas. 



```{r}
data_q3 <- country %>% 
  filter(Indicator.Code == "SP.DYN.TFRT.IN" | Indicator.Code == "NY.GDP.MKTP.CD" |  Indicator.Code == "SE.SCH.LIFE.FE" |  Indicator.Code == "SE.XPD.TOTL.GD.ZS" |  Indicator.Code == "SP.RUR.TOTL.FE.ZS")
```

```{r}
data_q3 <- data_q3 %>% mutate_if(is.factor, funs(case_when( 
    . == "SP.DYN.TFRT.IN" ~ "FertilityRate",
    . == "NY.GDP.MKTP.CD" ~ "GDP(log in billion)",
    . == "SE.SCH.LIFE.FE" ~ "ExpSchoolYear",
    . == "SE.XPD.TOTL.GD.ZS" ~ "EducationSpend",
    . == "SP.RUR.TOTL.FE.ZS" ~ "RuralPop",
    . == "CAN" ~ "North America", 
    . == "USA" ~ "North America",
    . == "MEX" ~ "North America",
    . == "NIC" ~ "North America",
    . == "SLV" ~ "North America",
    . == "COL" ~ "South America",
    . == "BRA" ~ "South America",
    . == "CHL" ~ "South America",
    . == "ARG" ~ "South America",
    . == "URY" ~ "South America",
    . == "RUS" ~ "Europe",
    . == "ISL" ~ "Europe",
    . == "UKR" ~ "Europe",
    . == "BLR" ~ "Europe",
    . == "LTU" ~ "Europe",
    . == "CHN" ~ "Asia",
    . == "JPN" ~ "Asia",
    . == "KOR" ~ "Asia",
    . == "THA" ~ "Asia",
    . == "MYS" ~ "Asia",
    . == "COG" ~ "Africa",
    . == "ZWE" ~ "Africa",
    . == "TZA" ~ "Africa",
    . == "NER" ~ "Africa",
    . == "NGA" ~ "Africa",
    . == "AUS" ~ "Oceania",
    . == "NZL" ~ "Oceania",
    . == "PNG" ~ "Oceania",
    . == "WSM" ~ "Oceania",
    . == "FJI" ~ "Oceania",
    T ~ as.character(.))))
```


```{r}
df <- data_q3 %>% 
  select(ï..Country.Name,Country.Code, Indicator.Code, c(40:64)) %>%
  group_by(ï..Country.Name, Country.Code, Indicator.Code)%>%
  summarise_if(is.numeric, mean, na.rm = TRUE)
```



```{r}
value <- rowMeans(df[4:28],  na.rm = TRUE)
df$value <- value
```

```{r}
df <- dcast(df, ï..Country.Name + Country.Code ~ Indicator.Code)
colnames(df)[colnames(df) == "Country.Code"] <- "Continent"
```

```{r}
df <- df[ , c(1, 2, 6, 5, 3, 7, 4)] 
```


```{r}
df[,3] <- log(df[,3]/1000000000)
```

```{r}
dfj<-df[!(df$ï..Country.Name=="Japan" | df$ï..Country.Name == "Nigeria"),]
```


```{r, fig.width = 12, fig.height = 12}
parcoords(dfj[2:7],
          rownames = F,
          brushMode = "1D-axes",
          reorderable = T,
          queue = T,
          alpha = 0.5,
          color = list(
            colorBy = "Continent",
            colorScale = "scaleOrdinal",
            colorScheme = c("#0c2d98", "#78411f", "#df5f5b", "#2b74e2", 
                            "#3bccf6", "#f4cf60")),
          withD3 = TRUE,
          autoresize = TRUE
          )
```

In this plot, we mainly focused on the factors related to women social status. The above parallel coordinates plot shows the relation between 6 parameters by using the data from 30 countries, 5 in each continent, after taking average of recent 25 years. These parameters are continent, GDP, fertility rate, government expenditure on education (% of GDP), female rural population (percentage) and expected schooling year for female. 

Here are some observations from this plot:

1. There are big differences in GDP even with countries in the same continent. However, in our sample, North America and Asia have relative higher GDP, whereas Africa and Oceania have relative lower values.

2. There is a clear inverse relationship between GDP and fertility rate: Countries with higher GDP have lower fertility rate. All 5 sample countries from Africa get high fertility rates (approximately 4.5 to 7.3). Oceania is the continent with second highest fertility rate. Asia and Europe have low fertility rate for below 2. 

3. Fertility rate and government expenditure on education are negatively correlated. 

4. Government expenditure on education is negatively correlated with female rural population.

5. Female rural population and female expected year of schooling has inverse relationship.

From these observations, we can identify there is a correlation between these variables. Women education status is correlated with geography, national wealth, fertility rate, rural population and government expenditure on education. For countries with higher income, the education expenditure is also likely to be higher as the data is percentage based on the total GDP. For countries with relative low GDP, especially in Africa, the fertility rates and rural population are both tend to be higher. In this case, the education resources are limited by both the economic factor and the children population. Moreover, women with higher educational level are likely to delay the age of getting marriage and giving birth, this can also explained by the correlation between low fertility rate and high expected female school year.

## Employment
```{r, echo = FALSE, fig.height = 10, fig.width = 10}
#Vincent Figure 1

packages = c('tidyverse', 'dplyr', 'tidyr')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(tidyr)

#Load in data

setwd("D:/Columbia University/GR5702 - EDAV/Final Project/Gender_Stats_csv")

genderStatsCountry <- read.csv("Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("Gender_StatsData.csv")
genderStatsFootnote <- read.csv("Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV',
              'COL', 'BRA', 'CHL', 'ARG', 'URY',
              'RUS', 'ISL', 'UKR', 'BLR', 'LTU',
              'CHN', 'JPN', 'KOR', 'THA', 'MYS',
              'ZWE', 'TZA', 'NER', 'NGA', 'COG',
              'AUS', 'NZL', 'PNG', 'WSM', 'FJI')

#Processing functions
capFirst <- function(s) {
  paste(toupper(substring(s, 1, 1)), substring(s, 2), sep = "")
}

lowFirst <- function(s) {
  paste(tolower(substring(s, 1, 1)), substring(s, 2), sep = "")
}

#Employment to population ratio by sex
emp_data <- full_data %>% 
  filter(Indicator.Name.x == 'Employment to population ratio, 15+, female (%) (modeled ILO estimate)' |
           Indicator.Name.x == 'Employment to population ratio, 15+, male (%) (modeled ILO estimate)') %>% 
  select(22:24, 75:84) %>% 
  mutate_if(is.numeric, round, 1)
emp_data <- emp_data[emp_data$Country.Code %in% countries ,]
emp_data <- emp_data %>% 
  mutate(emp_data, Mean = rowMeans(select(., starts_with('X')), na.rm = TRUE)) %>% 
  select(1:3, 14) %>% 
  separate(Indicator.Name.y, c('Indicator.Name', 'Age', 'Gender'), sep = ', ') %>% 
  separate(Gender, c('Sex', 'Other'), sep = '^[^\\s]*\\K\\s') %>% 
  select(1:3, 7, 5)
emp_data$Sex <- capFirst(emp_data$Sex)

#Plot Cleveland of Employment to Population Ratio
theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(1.25)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(1.25)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 1),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#D55E00", "#0072B2", "#F0E442", "#CC79A7")

#fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),

ggplot(emp_data, aes(Mean,
                     fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),
                     color = Sex)) +
  geom_point(size = 3) +
  ggtitle("2010-2019 Average Employment Ratio for each Country by Sex") + 
  ylab("") +
  xlab("% Employed") + 
  scale_colour_manual(values = cbPalette) +
  theme_dotplot
```
The figure above for the average employment ratio by sex between 2010-2019 is ordered in descending order by male employment ratio. The purpose of this ordering is to list countries with high male employment ratios first and visualize whether or not the female employment ratios match up. Out of the 30 countries selected for the study, it is very telling that a lower female employment ratio fairly universal. The actual differences between male/female employment ratio varies drastically by country. From the figure above, it is clear that the issue is much more pronounced in countries like Samoa, Mexico, and Malaysia (among others) as opposed to countries like Canada, Congo, Rep., and Iceland. Whether or not the differences between the ratio of male/female employment varies by continent is explored in another figure below. 
```{r, echo = FALSE, fig.height = 10, fig.width = 10}
#Vincent Figure 2

#Employment in upper/middle management 
mng_data <- full_data %>% 
  filter(Indicator.Name.x == 'Female share of employment in senior and middle management (%)') %>%
  select(22:24, 75:84) %>%
  mutate_if(is.numeric, round, 1)
mng_data <- mng_data[mng_data$Country.Code %in% countries ,]
mng_data <- mng_data %>% 
  mutate(mng_data, Mean = rowMeans(select(., starts_with('X')), na.rm = TRUE)) %>% 
  select(1:3, 14) %>% 
  separate(Indicator.Name.y, c('Sex', 'Indicator.Name'), sep = ' share of ')

male_percentage <- as.data.frame(100 - mng_data$Mean)
male_percentage <- male_percentage %>% 
  rename(Mean = '100 - mng_data$Mean')
male_vals <- cbind(ï..Country.Name = mng_data$ï..Country.Name, 
                    Country.Code = mng_data$Country.Code,
                    'Sex' = 'male',
                    Indicator.Name = mng_data$Indicator.Name,
                    Mean = male_percentage)

mng_data <- rbind(mng_data, male_vals) %>% 
  select(1:2, 4:5, 3)
mng_data$Indicator.Name <- capFirst(mng_data$Indicator.Name)
mng_data$Sex <- capFirst(mng_data$Sex)
mng_data$Indicator.Name <- substr(mng_data$Indicator.Name, 1, 
                                         nchar(mng_data$Indicator.Name) - 4)
mng_data$Mean <- as.numeric(mng_data$Mean)
mng_data[is.na(mng_data)] <- 0

#Plot Cleveland of Management Data
ggplot(mng_data, aes(Mean,
                     fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),
                     color = Sex)) +
  geom_point(size = 3) +
  ggtitle("2010-2019 Average Proportion of Upper/Middle Management for each Country by Sex") + 
  ylab("") +
  xlab("Proportion") + 
  xlim(10, 90) +
  scale_colour_manual(values = cbPalette) +
  theme_dotplot

```
The figure above for the average proportion of upper/middle management by sex is ordered in descending order by male upper/middle managers. The purpose of this ordering is to list countries with the biggest discrepancy in male/female manager ratio first, which can serve as a signal for gender discrepancies in employment opportunities. The countries on the figure without data points are countries where there is no data related to this area. These missing countries were kept within the figure to signify the fact that the International Labor Organization was unable to obtain any data related to this issue from numerous countries over a 10-year span. Even without 12 of the 30 randomly selected nations from the study, there appears to be a trend that people in upper/middle management are male the vast majority of the time in almost all countries with the exception of Nicaragua (in the sample). An interesting note is that although Papua New Guinea has a near equal proportion of female employment to male employment, upper/middle managers in that country are over 75% male. On the other hand, Nicaragua has a much higher male employment ratio than female employment ratio, but most upper/middle managers are female.
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 3

packages = c('tidyverse', 'dplyr', 'grid', 'gridExtra')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)

#Load in data

setwd("D:/Columbia University/GR5702 - EDAV/Final Project/Gender_Stats_csv")

genderStatsCountry <- read.csv("Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("Gender_StatsData.csv")
genderStatsFootnote <- read.csv("Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

#Full Data of all Countries
data <- full_data %>% 
  filter(Indicator.Name.x == 'Women Business and the Law Index Score (scale 1-100)') %>% 
  select(22:23, 55:84) %>% 
  mutate_if(is.numeric, round, 1)

#NA Region Data
na_countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV')
na_data <- data[data$Country.Code %in% na_countries ,]

names(na_data) <- substring(names(na_data), 2)
na_data <- na_data %>% 
  rename(Country.Code = 'ountry.Code')

na_data <- na_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

na_data$region = 'North America'
  
#SA Region Data
sa_countries = c('COL', 'BRA', 'CHL', 'ARG', 'URY')
sa_data <- data[data$Country.Code %in% sa_countries ,]

names(sa_data) <- substring(names(sa_data), 2)
sa_data <- sa_data %>% 
  rename(Country.Code = 'ountry.Code')

sa_data <- sa_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

sa_data$region = 'South America'

#EU Region Data
eu_countries = c('RUS', 'ISL', 'UKR', 'BLR', 'LTU')
eu_data <- data[data$Country.Code %in% eu_countries ,]

names(eu_data) <- substring(names(eu_data), 2)
eu_data <- eu_data %>% 
  rename(Country.Code = 'ountry.Code')

eu_data <- eu_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

eu_data$region = 'Europe'

#AS Region Data 
as_countries = c('CHN', 'JPN', 'KOR', 'THA', 'MYS')
as_data <- data[data$Country.Code %in% as_countries ,]

names(as_data) <- substring(names(as_data), 2)
as_data <- as_data %>% 
  rename(Country.Code = 'ountry.Code')

as_data <- as_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

as_data$region = 'Asia'

#AF Region Data 
af_countries = c('ZWE', 'TZA', 'NER', 'NGA', 'COG')
af_data <- data[data$Country.Code %in% af_countries ,]

names(af_data) <- substring(names(af_data), 2)
af_data <- af_data %>% 
  rename(Country.Code = 'ountry.Code')

af_data <- af_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

af_data$region = 'Africa'

#OC Region Data
oc_countries = c('AUS', 'NZL', 'PNG', 'WSM', 'FJI')
oc_data <- data[data$Country.Code %in% oc_countries ,]

names(oc_data) <- substring(names(oc_data), 2)
oc_data <- oc_data %>% 
  rename(Country.Code = 'ountry.Code')

oc_data <- oc_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

oc_data$region = 'Oceania'


#Plotting
full_data <- rbind(na_data, sa_data, eu_data, as_data, af_data, oc_data)

df <- split(full_data, f = full_data$region)

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#D55E00", "#0072B2", "#F0E442", "#CC79A7")

na <- ggplot(data = df$'North America', 
             aes(x = year, y = index_score, color = ..Country.Name, group = Country.Code)) + 
  geom_point() +
  geom_line() +
  facet_wrap(~region, ncol = 1) +
  labs(x = 'Year', y = 'Index Score') +
  theme_grey(16) +
  ylim(30, 100) +
  scale_colour_manual(values = cbPalette) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1),
        legend.title = element_blank())

sa <- na %+% df$'South America'
eu <- na %+% df$'Europe'
as <- na %+% df$'Asia'
af <- na %+% df$'Africa'
oc <- na %+% df$'Oceania'

grid.arrange(na,
             sa,
             eu,
             as,
             af,
             oc,
             top = textGrob("Women Business and the Law Index Score by Region Over the Years",
                            gp=gpar(fontsize=20,font=3)))

```
The Women, Business and the Law Score is an index that analyzes laws and regulations that affect women’s economic inclusion in 190 economies. It is composed of eight indicators revolving around women’s interactions with the law as they progress through their careers and make economic decisions. The eight indicators are: mobility, workplace, pay, marriage, parenthood, entrepreneurship, assets, and pension. The index ranges from 1-100 with 100 being the highest score, or equal legal rights with men in measured areas.  
  
The figure above displays the Women, Business and the Law index scores over a 30-year period faceted by continent with 5 countries per continent. The figure indicates a universal trend upwards in the index score over nearly all countries in every continent, although the degree of improvement varies drastically. For example, countries like the Congo, Rep. saw no improvement since 1900 and Papua New Guinea saw very marginal improvements, whereas countries like Brazil and Lithuania saw drastic improvements over the same time frame. It is very uncommon for the index score to go downwards once it has gone up, although it has happened on occasions like in 2004 for Lithuania, 2014 in Ukraine, and 2008 in China. A significant marker for gender inequality in employment is that only 2 of the 30 selected countries ever achieved a score of 100, signifying equal legal rights with men in the measured areas of the index. Although women’s legal rights in the eight indicator areas have improved worldwide since 1990, there are very few countries where equality is achieved as of now. 

```{r, echo = FALSE, fig.height = 10, fig.width = 10}
#Vincent Figure 4

packages = c('tidyverse', 'dplyr', 'grid', 'gridExtra', 'scales', 'patchwork')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)
library(scales)
library(patchwork)

#Load in data

setwd("D:/Columbia University/GR5702 - EDAV/Final Project/Gender_Stats_csv")

data = read.csv("Gender_StatsData.csv",header = TRUE,encoding="UTF-8")
colnames(data)=c("Country","Code","Indicator_details","Indicator",as.character(1960:2020))

Code = c("CAN","USA","MEX","NIC","SLV","COL","BRA","CHL","ARG","URY","RUS","ISL","UKR","BLR","LTU","CHN","JPN","KOR","THA","MYS","COG","ZWE","TZA","NER","NGA",'AUS','NZL','PNG','WSM','FJI')
Continent = rep(c("North America","South America","Europe","Asia","Africa","Oceania"),rep(5,6))
data_continent = data.frame(cbind(Code,Continent))

data_GDP = data[data$Code%in%Code & data$Indicator== "NY.GDP.PCAP.CD",c(2,64)]
colnames(data_GDP) = c("Code","GDP")
data_GDP$GDP = log(data_GDP$GDP)


data_incom = read.csv("Gender_StatsCountry.csv",header = TRUE,encoding="UTF-8")[,c(1,9)]
colnames(data_incom) = c("Code","Income")
data_INC = data_incom[data_incom$Code%in%Code,]
data_EMP = data[data$Code%in%Code & data$Indicator == "SL.TLF.CACT.FM.ZS",c(2,55:64)]
colnames(data_EMP) = c("Code",as.character(c(2010:2019)))
data_EMP = data_EMP%>%pivot_longer(cols = starts_with("1") | starts_with("2"),names_to = "Year",values_to = "ratio")
data_EMP = data_EMP%>% group_by(Code) %>% summarise(Mean = sum(ratio)/n())
data_EMP_AN = left_join(left_join(left_join(data_EMP,data_GDP,by=c("Code")),data_continent,by=c("Code")),data_INC,by = c("Code"))

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#0072B2", "#CC79A7")

ggplot(data_EMP_AN,aes(y=Mean,x=GDP, color = Continent))+
  geom_point(size = 3) +
  ggtitle('2010-2019 Average Female/Male Labor Force Participation Rate vs. Log-Scale GDP') +
  scale_colour_manual(values = cbPalette) +
  scale_x_continuous("Log(GDP)") +
  scale_y_continuous("Average of Female/Male Labor Force Participation Rate") +
  theme_grey(16) +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

```
The figure above displays the ratio of female to male employment ratio vs. GDP on a log scale for each country color coded by continent. There appears to no correlation overall between a country’s female/male labor force participation rate to the country’s GDP as the data points are scattered throughout as opposed to group along a linear line. However, when looking at individual continents separately, female/male labor force participation rate does seem correlated with GDP in continents like Europe, Africa, and North America (trend is not really observable for Asia, Oceania, and South America). This suggests that the correlation between female/male employment ratio may have a lot to do with geographical location of the country, and depending on the location, the ratio may be correlated with GDP.

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component

This interactive part shows the general relationship and the general trend of education, employment and social rights of women in different countries during last 25 years.

The social rights used the values of average score of 6 aspects mentioned above (showed in color, with lightgreen representing smaller values and black representing larger values). Darker circles represent higher equality in social rights. The employment used the values of employment ratio in percentage (showed in x axis). The education used the primary school enrollment rate of women (%)(showed in y axis).

Each point represents one country we sampled.

Users can click the light-green buttons which represents the year range of the statistics and see how the statistics for each country change and whether the different factors change in the similar direction.

This interactive portion shows the animation of changes that have happened in the past 25 years regarding social rights, education, and employment for all 30 countries simultaneously, something a static plot will struggle to present.

<script src="https://d3js.org/d3.v6.js"></script>
<svg width="688" height="500"></svg>

<script> 

  const svg = d3.select('svg');
  const width = +svg.attr('width');
  const height = +svg.attr('height');

  const render = data_1 => {
    const title = 'Social, Educational, and Employment Trends';
    const xValue_1 = d => d.emp1;
    const yValue_1 = d => d.educ1;
    const xValue_2 = d => d.emp2;
    const yValue_2 = d => d.educ2;
    const xValue_3 = d => d.emp3;
    const yValue_3 = d => d.educ3;
    const xValue_4 = d => d.emp4;
    const yValue_4 = d => d.educ4;
    const xValue_5 = d => d.emp5;
    const yValue_5 = d => d.educ5;
    const rValue_1 = d => d.Score1;
    const rValue_2 = d => d.Score2;
    const rValue_3 = d => d.Score3;
    const rValue_4 = d => d.Score4;
    const rValue_5 = d => d.Score5;
    const xAxisLabel = 'Employment';
    const yAxisLabel = 'Education';
    const circleRadius = 10;
    const margin = { top: 60, right: 40, bottom: 175, left: 150 };
    const innerWidth = width - margin.left - margin.right;
    const innerHeight = height - margin.top - margin.bottom;
    const xScale = d3.scaleLinear()
      .domain([40, 100])
      .range([0, innerWidth])
      .nice();
    const yScale = d3.scaleLinear()
      .domain([20, 180])
      .range([innerHeight, 0])
      .nice();
    const compute = d3.interpolate('lightgreen', 'black')
    const linear = d3.scaleLinear().domain([0.46, 1]).range([0, 1])
  	const myColor = d3.scaleLinear().domain([0.3, 1])
    	.range(["black", "green"]);
    const g = svg.append('g')
      .attr('transform', `translate(${margin.left},${margin.top})`);
    const xAxis = d3.axisBottom(xScale)
      .tickSize(-innerHeight)
      .tickPadding(15);
    const yAxis = d3.axisLeft(yScale)
      .tickSize(-innerWidth)
      .tickPadding(10);
    const yAxisG = g.append('g').call(yAxis);
    yAxisG.selectAll('.domain').remove();
    yAxisG.append('text')
        .attr('class', 'axis-label')
        .attr('y', -93)
        .attr('x', -innerHeight / 2)
        .attr('fill', 'black')
        .attr('transform', `rotate(-90)`)
        .attr('text-anchor', 'middle')
        .text(yAxisLabel);
    const xAxisG = g.append('g').call(xAxis)
      .attr('transform', `translate(0,${innerHeight})`);
    xAxisG.select('.domain').remove();
    xAxisG.append('text')
        .attr('class', 'axis-label')
        .attr('y', 75)
        .attr('x', innerWidth / 2)
        .attr('fill', 'black')
        .text(xAxisLabel);
    g.selectAll('circle').data(data_1)
      .enter().append('circle')
        .attr('cy', d => yScale(yValue_1(d)))
        .attr('cx', d => xScale(xValue_1(d)))
        .attr('r',  d => circleRadius)
    		.attr('fill', d => compute(linear(rValue_1(d))));
    g.append('text')
        .attr('class', 'title')
        .attr('y', -10)
        .text(title);

  var object_dataset = [
    {cx: 43, cy: 450, text: `1995-1999`,no:"r1"},
    {cx: 172, cy: 450, text: `2000-2004`,no:"r2"},
    {cx: 301, cy: 450, text: `2005-2009`,no:"r3"},
    {cx: 430, cy: 450, text: `2010-2014`,no:"r4"},
    {cx: 559, cy: 450, text: `2015-2019`,no:"r5"},
    ];

  svg.selectAll('rect').data(object_dataset)
    .enter().append('rect')
      .attr('x', d => d.cx + 50)
      .attr('y', d => d.cy - 30)
      .attr('width', 86)
      .attr('height', 50)
      .attr('fill', 'lightgreen')
      .attr('id', d => d.no);

  svg.select("rect#r2").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r2").attr("fill","green")
        svg.selectAll("circle").transition()
        .duration(2000)
        .attr('cy', d => yScale(yValue_2(d)))
        .attr('cx', d => xScale(xValue_2(d)))
        .attr('r',  d => circleRadius)
        .attr('fill', "red")
        .attr('fill', d => compute(linear(rValue_2(d))))})
      ;
          
  svg.select("rect#r1").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r1").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_1(d)))
    .attr('cx', d => xScale(xValue_1(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_1(d))))})
  ;
          
  svg.select("rect#r3").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r3").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_3(d)))
    .attr('cx', d => xScale(xValue_3(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_3(d))))})
  ;

  svg.select("rect#r4").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r4").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_4(d)))
    .attr('cx', d => xScale(xValue_4(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_4(d))))})
  ;
  
  svg.select("rect#r5").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r5").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_5(d)))
    .attr('cx', d => xScale(xValue_5(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_5(d))))})
  };

  d3.csv('https://raw.githubusercontent.com/yaojiaye00/final5702/main/data_all.csv')
    .then(data_1 => {
      data_1.forEach(d => {
        d.Score1 = +d.Score1;
        d.educ1 = +d.educ1; //Currently need to manually update based on input 
        d.emp1 = +d.emp1; //Same as above
        d.Score2 = +d.Score2;
        d.educ2 = +d.educ2; 
        d.emp2 = +d.emp2; 
        d.Score3 = +d.Score3;
        d.educ3 = +d.educ3; 
        d.emp3 = +d.emp3; 
        d.Score4 = +d.Score4;
        d.educ4 = +d.educ4; 
        d.emp4 = +d.emp4; 
        d.Score5 = +d.Score5;
        d.educ5 = +d.educ5; 
        d.emp5 = +d.emp5; 
      });
      render(data_1)
    });

  var object_dataset = [
    {cx: 43, cy: 450, text: `1995-1999`,no:"r1"},
    {cx: 172, cy: 450, text: `2000-2004`,no:"r2"},
    {cx: 301, cy: 450, text: `2005-2009`,no:"r3"},
    {cx: 430, cy: 450, text: `2010-2014`,no:"r4"},
    {cx: 559, cy: 450, text: `2015-2019`,no:"r5"},
    ];

  svg.selectAll('text').data(object_dataset)
    .enter().append('text')
      .attr('x', d => d.cx + 90)
      .attr('y', d => d.cy + 50)
      .text(d => d.text)
      .attr('fill', 'black')
      .attr('text-anchor', 'middle');

</script>


The interactive plot above was built off a format provided by "Cars Scatter Plot" example from Vizhub (https://vizhub.com/curran/9247d4d42df74185980f7b1f7504dcc5).  
  


<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion

For social rights:    
1.In general, Iceland, Canada and New Zealand do well. No obvious gender discrimination in finance and freedom, but marked gender discrimination is found in social welfare these year.  
2.World started to focus on solving the gender discrimination problem at around 1987. The gender discrimination decreased a lot during last 40 years. And in the law and welfare aspects, the discrimination moderated rapidly yet still exists.  
3.GDP has no influence on the gender discrimination in social rights aspects while the location of the country will, to some extent, determine the gender discrimination condition in social rights aspect. Generally, the North America will have a higher scores in the majority of topics.  
  
For education:  
1. There exists gender discrimination on education opportunities. Even in recent years, share of female graduates from different majors still have big differences.  
2. The education level of women is correlated with both GDP and geographical factor. Countries with high GDP will highly likely to have more education resources and less children in rural area where education resources are more scarce. Countries in North and South America have high scores in female education level shows there may also geographical factor that correlates with gender discrimination on education.  
  
For employment:  
1) There appears to be noticeable employment differences between genders in almost all of the countries sampled for the study. The differences observed stemmed from employment ratios and potentially career opportunities. There was only one case where a country had more females in upper/middle management than males, the others were all heavily skewed towards male.  
2) over the past 30 years, the trends for gender equality in employment has been steadily getting better throughout the world, as suggested by growing Women Business and the Law index scores for every nation selected (though significant discrepancies between nations exist). As for regional or developmental effects on employment ratios, the trends appear unobservable or very weak at best.  
3) There appears to be more correlation between GDP and employment ratios throughout the world, but if you subset to specific continents, an argument can be made that specific continents like Europe and Africa have correlated GDP and female/male labor force participation rates.  

Some limitations:  
1.Owing to the availability of the data, we only sampled 30 countries from different continents in the world. The analysis result may not be representative for the whole world.  
2.The number of social rights questions asked by the dataset provider is not large enough (only 24), so the analysis result may be limited and only can reflect partial conditions of the worldwide social rights for women.  
3.Maybe when more data are available, we can explore more about the same topic (analysis limitation).  
4.The assignment of questions to topics is quite subjective. Maybe there are some better assignments for the questions to the topics.  

<!--chapter:end:07-conclusion.Rmd-->

