--- 
title: "Analyzing Gender Discrepancies Over Time Using Social, Educational, and Employment Trends"
author: "Vincent Chang, Xiaoyan Li, Yaojia Ye"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction


<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data sources

The data for this project is sourced from the World Bank. The World Bank is a group dedicated to providing high-quality statistical data on numerous topics globally to aid development and policy making. A lot of the information within the World Bank’s databases come directly from the statistical systems of member countries, which means the accuracy of the information is heavily dependent on how well these national systems perform their data collection. As a result, a high priority of the World Bank is to help improve statistical infrastructures of developing countries by preparing national strategies for the development of statistics “as recommended in the Marrakech Action Plan for Statistics”. Data collection is currently a joint effort between the World Bank and many other international statistical communities including the United Nations (UN), the Organization for Economic Co-Operation and Development (OECD), the International Monetary Fund (IMF), and regional development banks. Transnational data is also collected by the World Bank through globally coordinated programs like the Marrakech Action Plan for Statistics and PARIS21.  

For all the data collected, the World Bank ensures data consistency and reliability through:  
  -	Participation in the UN Statistical Commission and other statistical forums to develop      appropriate frameworks, guidance, and standards of good practice for statistics  
  -	Building consensus and defining internationally agreed indicators such as the indicators for the Millennium Development Goals  
  -	Establishing data exchange processes and methods  
  -	Assembling, analyzing, and disseminating data online and in print  
  
*Info comes from: https://data.worldbank.org/about

<!--chapter:end:02-data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Data transformation

Data for this project is sourced from the following link:   https://datacatalog.worldbank.org/dataset/gender-statistics  
  
The CSV files were downloaded from the "Data & Resources" tab and imported directly into RStudio for processing. The specific steps taken to process the data for visualization differed between the three topics studied for this project: social rights, education, and employment.  

For social rights,  
  
  
For education,  
  
  
For employment, 

<!--chapter:end:03-cleaning.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Missing values

<!--chapter:end:04-missing.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

## Social

## Education

## Employment
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 1

packages = c('tidyverse', 'dplyr', 'tidyr')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(tidyr)

#Load in data

genderStatsCountry <- read.csv("data/Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("data/Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("data/Gender_StatsData.csv")
genderStatsFootnote <- read.csv("data/Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("data/Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("data/Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV',
              'COL', 'BRA', 'CHL', 'ARG', 'URY',
              'RUS', 'ISL', 'UKR', 'BLR', 'LTU',
              'CHN', 'JPN', 'KOR', 'THA', 'MYS',
              'ZWE', 'TZA', 'NER', 'NGA', 'COG',
              'AUS', 'NZL', 'PNG', 'WSM', 'FJI')

#Processing functions
capFirst <- function(s) {
  paste(toupper(substring(s, 1, 1)), substring(s, 2), sep = "")
}

lowFirst <- function(s) {
  paste(tolower(substring(s, 1, 1)), substring(s, 2), sep = "")
}

#Employment to population ratio by sex
emp_data <- full_data %>% 
  filter(Indicator.Name.x == 'Employment to population ratio, 15+, female (%) (modeled ILO estimate)' |
           Indicator.Name.x == 'Employment to population ratio, 15+, male (%) (modeled ILO estimate)') %>% 
  select(22:24, 75:84) %>% 
  mutate_if(is.numeric, round, 1)
emp_data <- emp_data[emp_data$Country.Code %in% countries ,]
emp_data <- emp_data %>% 
  mutate(emp_data, Mean = rowMeans(select(., starts_with('X')), na.rm = TRUE)) %>% 
  select(1:3, 14) %>% 
  separate(Indicator.Name.y, c('Indicator.Name', 'Age', 'Gender'), sep = ', ') %>% 
  separate(Gender, c('Sex', 'Other'), sep = '^[^\\s]*\\K\\s') %>% 
  select(1:3, 7, 5)
emp_data$Sex <- capFirst(emp_data$Sex)

#Plot Cleveland of Employment to Population Ratio
theme_dotplot <- theme_bw(14) +
  theme(axis.text.y = element_text(size = rel(1.25)),
        axis.ticks.y = element_blank(),
        axis.title.x = element_text(size = rel(1.25)),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(size = 1),
        panel.grid.minor.x = element_blank(),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#D55E00", "#0072B2", "#F0E442", "#CC79A7")

#fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),

ggplot(emp_data, aes(Mean,
                     fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),
                     color = Sex)) +
  geom_point(size = 3) +
  ggtitle("2010-2019 Average Employment Ratio for each Country by Sex") + 
  ylab("") +
  xlab("% Employed") + 
  scale_colour_manual(values = cbPalette) +
  theme_dotplot
```
The figure above for the average employment ratio by sex between 2010-2019 is ordered in descending order by male employment ratio. The purpose of this ordering is to list countries with high male employment ratios first and visualize whether or not the female employment ratios match up. Out of the 30 countries selected for the study, it is very telling that a lower female employment ratio fairly universal. The actual differences between male/female employment ratio varies drastically by country. From the figure above, it is clear that the issue is much more pronounced in countries like Samoa, Mexico, and Malaysia (among others) as opposed to countries like Canada, Congo, Rep., and Iceland. Whether or not the differences between the ratio of male/female employment varies by continent is explored in another figure below. 
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 2

#Employment in upper/middle management 
mng_data <- full_data %>% 
  filter(Indicator.Name.x == 'Female share of employment in senior and middle management (%)') %>%
  select(22:24, 75:84) %>%
  mutate_if(is.numeric, round, 1)
mng_data <- mng_data[mng_data$Country.Code %in% countries ,]
mng_data <- mng_data %>% 
  mutate(mng_data, Mean = rowMeans(select(., starts_with('X')), na.rm = TRUE)) %>% 
  select(1:3, 14) %>% 
  separate(Indicator.Name.y, c('Sex', 'Indicator.Name'), sep = ' share of ')

male_percentage <- as.data.frame(100 - mng_data$Mean)
male_percentage <- male_percentage %>% 
  rename(Mean = '100 - mng_data$Mean')
male_vals <- cbind(ï..Country.Name = mng_data$ï..Country.Name, 
                    Country.Code = mng_data$Country.Code,
                    'Sex' = 'male',
                    Indicator.Name = mng_data$Indicator.Name,
                    Mean = male_percentage)

mng_data <- rbind(mng_data, male_vals) %>% 
  select(1:2, 4:5, 3)
mng_data$Indicator.Name <- capFirst(mng_data$Indicator.Name)
mng_data$Sex <- capFirst(mng_data$Sex)
mng_data$Indicator.Name <- substr(mng_data$Indicator.Name, 1, 
                                         nchar(mng_data$Indicator.Name) - 4)
mng_data$Mean <- as.numeric(mng_data$Mean)
mng_data[is.na(mng_data)] <- 0

#Plot Cleveland of Management Data
ggplot(mng_data, aes(Mean,
                     fct_reorder2(`ï..Country.Name`, Sex == "Male", Mean, .desc = FALSE),
                     color = Sex)) +
  geom_point(size = 3) +
  ggtitle("2010-2019 Average Proportion of Upper/Middle Management for each Country by Sex") + 
  ylab("") +
  xlab("Proportion") + 
  xlim(10, 90) +
  scale_colour_manual(values = cbPalette) +
  theme_dotplot

```
The figure above for the average proportion of upper/middle management by sex is ordered in descending order by male upper/middle managers. The purpose of this ordering is to list countries with the biggest discrepancy in male/female manager ratio first, which can serve as a signal for gender discrepancies in employment opportunities. The countries on the figure without data points are countries where there is no data related to this area. These missing countries were kept within the figure to signify the fact that the International Labor Organization was unable to obtain any data related to this issue from numerous countries over a 10-year span. Even without 12 of the 30 randomly selected nations from the study, there appears to be a trend that people in upper/middle management are male the vast majority of the time in almost all countries with the exception of Nicaragua (in the sample). An interesting note is that although Papua New Guinea has a near equal proportion of female employment to male employment, upper/middle managers in that country are over 75% male. On the other hand, Nicaragua has a much higher male employment ratio than female employment ratio, but most upper/middle managers are female.
```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 3

packages = c('tidyverse', 'dplyr', 'grid', 'gridExtra')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)

#Load in data

genderStatsCountry <- read.csv("data/Gender_StatsCountry.csv")
genderStatsCountrySeries <- read.csv("data/Gender_StatsCountry-Series.csv")
genderStatsData <- read.csv("data/Gender_StatsData.csv")
genderStatsFootnote <- read.csv("data/Gender_StatsFootNote.csv")
genderStatsSeries <- read.csv("data/Gender_StatsSeries.csv")
genderStatsSeriesTime <- read.csv("data/Gender_StatsSeries-Time.csv")
full_data <- merge(genderStatsSeries, genderStatsData, by.x = "ï..Series.Code", by.y = "Indicator.Code")

#Full Data of all Countries
data <- full_data %>% 
  filter(Indicator.Name.x == 'Women Business and the Law Index Score (scale 1-100)') %>% 
  select(22:23, 55:84) %>% 
  mutate_if(is.numeric, round, 1)

#NA Region Data
na_countries = c('CAN', 'USA', 'MEX', 'NIC', 'SLV')
na_data <- data[data$Country.Code %in% na_countries ,]

names(na_data) <- substring(names(na_data), 2)
na_data <- na_data %>% 
  rename(Country.Code = 'ountry.Code')

na_data <- na_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

na_data$region = 'North America'
  
#SA Region Data
sa_countries = c('COL', 'BRA', 'CHL', 'ARG', 'URY')
sa_data <- data[data$Country.Code %in% sa_countries ,]

names(sa_data) <- substring(names(sa_data), 2)
sa_data <- sa_data %>% 
  rename(Country.Code = 'ountry.Code')

sa_data <- sa_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

sa_data$region = 'South America'

#EU Region Data
eu_countries = c('RUS', 'ISL', 'UKR', 'BLR', 'LTU')
eu_data <- data[data$Country.Code %in% eu_countries ,]

names(eu_data) <- substring(names(eu_data), 2)
eu_data <- eu_data %>% 
  rename(Country.Code = 'ountry.Code')

eu_data <- eu_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

eu_data$region = 'Europe'

#AS Region Data 
as_countries = c('CHN', 'JPN', 'KOR', 'THA', 'MYS')
as_data <- data[data$Country.Code %in% as_countries ,]

names(as_data) <- substring(names(as_data), 2)
as_data <- as_data %>% 
  rename(Country.Code = 'ountry.Code')

as_data <- as_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

as_data$region = 'Asia'

#AF Region Data 
af_countries = c('ZWE', 'TZA', 'NER', 'NGA', 'COG')
af_data <- data[data$Country.Code %in% af_countries ,]

names(af_data) <- substring(names(af_data), 2)
af_data <- af_data %>% 
  rename(Country.Code = 'ountry.Code')

af_data <- af_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

af_data$region = 'Africa'

#OC Region Data
oc_countries = c('AUS', 'NZL', 'PNG', 'WSM', 'FJI')
oc_data <- data[data$Country.Code %in% oc_countries ,]

names(oc_data) <- substring(names(oc_data), 2)
oc_data <- oc_data %>% 
  rename(Country.Code = 'ountry.Code')

oc_data <- oc_data %>% 
  pivot_longer(cols = !Country.Code & !..Country.Name, names_to = 'year', values_to = 'index_score')

oc_data$region = 'Oceania'


#Plotting
full_data <- rbind(na_data, sa_data, eu_data, as_data, af_data, oc_data)

df <- split(full_data, f = full_data$region)

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#D55E00", "#0072B2", "#F0E442", "#CC79A7")

na <- ggplot(data = df$'North America', 
             aes(x = year, y = index_score, color = ..Country.Name, group = Country.Code)) + 
  geom_point() +
  geom_line() +
  facet_wrap(~region, ncol = 1) +
  labs(x = 'Year', y = 'Index Score') +
  theme_grey(16) +
  ylim(30, 100) +
  scale_colour_manual(values = cbPalette) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1),
        legend.title = element_blank())

sa <- na %+% df$'South America'
eu <- na %+% df$'Europe'
as <- na %+% df$'Asia'
af <- na %+% df$'Africa'
oc <- na %+% df$'Oceania'

grid.arrange(na,
             sa,
             eu,
             as,
             af,
             oc,
             top = textGrob("Women Business and the Law Index Score by Region Over the Years",
                            gp=gpar(fontsize=20,font=3)))

```
The Women, Business and the Law Score is an index that analyzes laws and regulations that affect women’s economic inclusion in 190 economies. It is composed of eight indicators revolving around women’s interactions with the law as they progress through their careers and make economic decisions. The eight indicators are: mobility, workplace, pay, marriage, parenthood, entrepreneurship, assets, and pension. The index ranges from 1-100 with 100 being the highest score, or equal legal rights with men in measured areas.  
  
The figure above displays the Women, Business and the Law index scores over a 30-year period faceted by continent with 5 countries per continent. The figure indicates a universal trend upwards in the index score over nearly all countries in every continent, although the degree of improvement varies drastically. For example, countries like the Congo, Rep. saw no improvement since 1900 and Papua New Guinea saw very marginal improvements, whereas countries like Brazil and Lithuania saw drastic improvements over the same time frame. It is very uncommon for the index score to go downwards once it has gone up, although it has happened on occasions like in 2004 for Lithuania, 2014 in Ukraine, and 2008 in China. A significant marker for gender inequality in employment is that only 2 of the 30 selected countries ever achieved a score of 100, signifying equal legal rights with men in the measured areas of the index. Although women’s legal rights in the eight indicator areas have improved worldwide since 1990, there are very few countries where equality is achieved as of now. 

```{r, echo = FALSE, fig.height = 15, fig.width = 15}
#Vincent Figure 4

packages = c('tidyverse', 'dplyr', 'grid', 'gridExtra', 'scales', 'patchwork')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
}

library(tidyverse)
library(dplyr)
library(grid)
library(gridExtra)
library(scales)
library(patchwork)

#Load in data


data = read.csv("data/Gender_StatsData.csv",header = TRUE,encoding="UTF-8")
colnames(data)=c("Country","Code","Indicator_details","Indicator",as.character(1960:2020))

Code = c("CAN","USA","MEX","NIC","SLV","COL","BRA","CHL","ARG","URY","RUS","ISL","UKR","BLR","LTU","CHN","JPN","KOR","THA","MYS","COG","ZWE","TZA","NER","NGA",'AUS','NZL','PNG','WSM','FJI')
Continent = rep(c("North America","South America","Europe","Asia","Africa","Oceania"),rep(5,6))
data_continent = data.frame(cbind(Code,Continent))

data_GDP = data[data$Code%in%Code & data$Indicator== "NY.GDP.PCAP.CD",c(2,64)]
colnames(data_GDP) = c("Code","GDP")
data_GDP$GDP = log(data_GDP$GDP)


data_incom = read.csv("data/Gender_StatsCountry.csv",header = TRUE,encoding="UTF-8")[,c(1,9)]
colnames(data_incom) = c("Code","Income")
data_INC = data_incom[data_incom$Code%in%Code,]
data_EMP = data[data$Code%in%Code & data$Indicator == "SL.TLF.CACT.FM.ZS",c(2,55:64)]
colnames(data_EMP) = c("Code",as.character(c(2010:2019)))
data_EMP = data_EMP%>%pivot_longer(cols = starts_with("1") | starts_with("2"),names_to = "Year",values_to = "ratio")
data_EMP = data_EMP%>% group_by(Code) %>% summarise(Mean = sum(ratio)/n())
data_EMP_AN = left_join(left_join(left_join(data_EMP,data_GDP,by=c("Code")),data_continent,by=c("Code")),data_INC,by = c("Code"))

cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#0072B2", "#CC79A7")

ggplot(data_EMP_AN,aes(y=Mean,x=GDP, color = Continent))+
  geom_point(size = 3) +
  ggtitle('2010-2019 Average Female/Male Labor Force Participation Rate vs. Log-Scale GDP') +
  scale_colour_manual(values = cbPalette) +
  scale_x_continuous("Log(GDP)") +
  scale_y_continuous("Average of Female/Male Labor Force Participation Rate") +
  theme_grey(16) +
  theme(legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5))

```
The figure above displays the ratio of female to male employment ratio vs. GDP on a log scale for each country color coded by continent. There appears to no correlation overall between a country’s female/male labor force participation rate to the country’s GDP as the data points are scattered throughout as opposed to group along a linear line. However, when looking at individual continents separately, female/male labor force participation rate does seem correlated with GDP in continents like Europe, Africa, and North America (trend is not really observable for Asia, Oceania, and South America). This suggests that the correlation between female/male employment ratio may have a lot to do with geographical location of the country, and depending on the location, the ratio may be correlated with GDP.

<!--chapter:end:05-results.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Interactive component

<script src="https://d3js.org/d3.v6.js"></script>
<svg width="688" height="500"></svg>

<script> 

  const svg = d3.select('svg');
  const width = +svg.attr('width');
  const height = +svg.attr('height');

  const render = data_1 => {
    const title = 'Social, Educational, and Employment Trends';
    const xValue_1 = d => d.emp1;
    const yValue_1 = d => d.educ1;
    const xValue_2 = d => d.emp2;
    const yValue_2 = d => d.educ2;
    const xValue_3 = d => d.emp3;
    const yValue_3 = d => d.educ3;
    const xValue_4 = d => d.emp4;
    const yValue_4 = d => d.educ4;
    const xValue_5 = d => d.emp5;
    const yValue_5 = d => d.educ5;
    const rValue_1 = d => d.Score1;
    const rValue_2 = d => d.Score2;
    const rValue_3 = d => d.Score3;
    const rValue_4 = d => d.Score4;
    const rValue_5 = d => d.Score5;
    const xAxisLabel = 'Employment';
    const yAxisLabel = 'Education';
    const circleRadius = 10;
    const margin = { top: 60, right: 40, bottom: 175, left: 150 };
    const innerWidth = width - margin.left - margin.right;
    const innerHeight = height - margin.top - margin.bottom;
    const xScale = d3.scaleLinear()
      .domain([40, 100])
      .range([0, innerWidth])
      .nice();
    const yScale = d3.scaleLinear()
      .domain([20, 180])
      .range([innerHeight, 0])
      .nice();
    const compute = d3.interpolate('lightgreen', 'black')
    const linear = d3.scaleLinear().domain([0.46, 1]).range([0, 1])
  	const myColor = d3.scaleLinear().domain([0.3, 1])
    	.range(["black", "green"]);
    const g = svg.append('g')
      .attr('transform', `translate(${margin.left},${margin.top})`);
    const xAxis = d3.axisBottom(xScale)
      .tickSize(-innerHeight)
      .tickPadding(15);
    const yAxis = d3.axisLeft(yScale)
      .tickSize(-innerWidth)
      .tickPadding(10);
    const yAxisG = g.append('g').call(yAxis);
    yAxisG.selectAll('.domain').remove();
    yAxisG.append('text')
        .attr('class', 'axis-label')
        .attr('y', -93)
        .attr('x', -innerHeight / 2)
        .attr('fill', 'black')
        .attr('transform', `rotate(-90)`)
        .attr('text-anchor', 'middle')
        .text(yAxisLabel);
    const xAxisG = g.append('g').call(xAxis)
      .attr('transform', `translate(0,${innerHeight})`);
    xAxisG.select('.domain').remove();
    xAxisG.append('text')
        .attr('class', 'axis-label')
        .attr('y', 75)
        .attr('x', innerWidth / 2)
        .attr('fill', 'black')
        .text(xAxisLabel);
    g.selectAll('circle').data(data_1)
      .enter().append('circle')
        .attr('cy', d => yScale(yValue_1(d)))
        .attr('cx', d => xScale(xValue_1(d)))
        .attr('r',  d => circleRadius)
    		.attr('fill', d => compute(linear(rValue_1(d))));
    g.append('text')
        .attr('class', 'title')
        .attr('y', -10)
        .text(title);

  var object_dataset = [
    {cx: 43, cy: 450, text: `1995-1999`,no:"r1"},
    {cx: 172, cy: 450, text: `2000-2004`,no:"r2"},
    {cx: 301, cy: 450, text: `2005-2009`,no:"r3"},
    {cx: 430, cy: 450, text: `2010-2014`,no:"r4"},
    {cx: 559, cy: 450, text: `2015-2019`,no:"r5"},
    ];

  svg.selectAll('rect').data(object_dataset)
    .enter().append('rect')
      .attr('x', d => d.cx + 50)
      .attr('y', d => d.cy - 30)
      .attr('width', 86)
      .attr('height', 50)
      .attr('fill', 'lightgreen')
      .attr('id', d => d.no);

  svg.select("rect#r2").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r2").attr("fill","green")
        svg.selectAll("circle").transition()
        .duration(2000)
        .attr('cy', d => yScale(yValue_2(d)))
        .attr('cx', d => xScale(xValue_2(d)))
        .attr('r',  d => circleRadius)
        .attr('fill', "red")
        .attr('fill', d => compute(linear(rValue_2(d))))})
      ;
          
  svg.select("rect#r1").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r1").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_1(d)))
    .attr('cx', d => xScale(xValue_1(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_1(d))))})
  ;
          
  svg.select("rect#r3").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r3").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_3(d)))
    .attr('cx', d => xScale(xValue_3(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_3(d))))})
  ;

  svg.select("rect#r4").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r4").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_4(d)))
    .attr('cx', d => xScale(xValue_4(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_4(d))))})
  ;
  
  svg.select("rect#r5").on("click",function(data_1){
    svg.selectAll("rect").attr("fill","lightgreen")
    svg.select("rect#r5").attr("fill","green")
    svg.selectAll("circle").transition()
    .duration(2000)
    .attr('cy', d => yScale(yValue_5(d)))
    .attr('cx', d => xScale(xValue_5(d)))
    .attr('r',  d => circleRadius)
    .attr('fill', "red")
    .attr('fill', d => compute(linear(rValue_5(d))))})
  };

  d3.csv('https://raw.githubusercontent.com/yaojiaye00/final5702/main/data_all.csv')
    .then(data_1 => {
      data_1.forEach(d => {
        d.Score1 = +d.Score1;
        d.educ1 = +d.educ1; //Currently need to manually update based on input 
        d.emp1 = +d.emp1; //Same as above
        d.Score2 = +d.Score2;
        d.educ2 = +d.educ2; 
        d.emp2 = +d.emp2; 
        d.Score3 = +d.Score3;
        d.educ3 = +d.educ3; 
        d.emp3 = +d.emp3; 
        d.Score4 = +d.Score4;
        d.educ4 = +d.educ4; 
        d.emp4 = +d.emp4; 
        d.Score5 = +d.Score5;
        d.educ5 = +d.educ5; 
        d.emp5 = +d.emp5; 
      });
      render(data_1)
    });

  var object_dataset = [
    {cx: 43, cy: 450, text: `1995-1999`,no:"r1"},
    {cx: 172, cy: 450, text: `2000-2004`,no:"r2"},
    {cx: 301, cy: 450, text: `2005-2009`,no:"r3"},
    {cx: 430, cy: 450, text: `2010-2014`,no:"r4"},
    {cx: 559, cy: 450, text: `2015-2019`,no:"r5"},
    ];

  svg.selectAll('text').data(object_dataset)
    .enter().append('text')
      .attr('x', d => d.cx + 90)
      .attr('y', d => d.cy + 50)
      .text(d => d.text)
      .attr('fill', 'black')
      .attr('text-anchor', 'middle');

</script>

<!--chapter:end:06-interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion


<!--chapter:end:07-conclusion.Rmd-->

